---
# ============================================================
# YAML HEADER
# This block controls the overall appearance and layout of the 
# dashboard. It is not R code — it is read by the flexdashboard 
# package when knitting the document.
# ============================================================

title: "VCFG Dataset 2: Cell Viability"   # Title shown in the navbar
output: 
  flexdashboard::flex_dashboard:
    theme:                    # Custom colour theme using Bootstrap Sass variables - if using a preseet Bootswatch theme, specify it here
      bg: "white"             # Background colour of the dashboard
      fg: "#3c2372"           # Foreground (text) colour — a dark purple
      primary: "#3c2372"      # Primary accent colour (buttons, active elements)
      secondary: "#3c2372"    # Secondary accent colour
      success: "#3c2372"      # Colour for "success" state elements
    orientation: columns      # Panels are arranged side-by-side in a column-based layout
    vertical_layout: fill     # Panels expand to fill the full height of the browser
---

<!-- ============================================================
  CSS STYLING BLOCK
  Raw HTML/CSS injected directly into the dashboard. This customises
  the appearance of specific elements beyond what the YAML theme covers.
  Each section targets a different part of the UI. Sections can be added
  or removed depending on your customisation requirements.
  ============================================================ -->

<style type="text/css">

/* ---------------------------------------------------------
   DataTables Buttons (Copy, Excel, Show/Hide Columns)
   Makes the row of export buttons compact and left-aligned,
   rather than stretching to fill the available width.
   --------------------------------------------------------- */

.dataTables_wrapper .dt-buttons {
  display: inline-flex !important;  /* Lay buttons out in a horizontal row */
  align-items: center;
  gap: 6px;                         /* Small gap between each button */
  width: auto !important;           /* Don't stretch to full container width */
  flex: 0 0 auto !important;
}

.dt-buttons .dt-button {
  font-size: 13px !important;
  padding: 4px 8px !important;
  background: #f8f8f8 !important;   /* Light grey button background */
  border: 1px solid #bbb !important;
  border-radius: 4px !important;
}

.dt-buttons .dt-button:hover {
  background: #eaeaea !important;   /* Slightly darker on hover */
  border-color: #999 !important;
}

/* ---------------------------------------------------------
   Chart Panel Titles
   The text shown above each panel in the dashboard.
   --------------------------------------------------------- */

.chart-title {  
  font-size: 19px;
  font-weight: bold;
  color: #3c2372;   /* Match the dashboard's purple theme colour */
}

/* ---------------------------------------------------------
   ionRangeSlider (the interactive slider filter widgets)
   The slider is created by the crosstalk filter_slider() function.
   These rules customise its colours and hide default labels that
   would otherwise clutter the UI.
   --------------------------------------------------------- */

.irs-grid-text {
  color: #3c2372;
  display: none !important;   /* Hide the numeric grid labels beneath the slider */
}

.irs-bar {
  background-color: #3c2372; /* Purple fill for the selected range */
  border: 1px solid #3c2372;
}

.irs-slider {
  background: #3c2372 !important;   /* The draggable handle colour */
  border: 2px solid #3c2372;
  width: 10px !important;
  height: 10px !important;
  top: 25px !important;             /* Vertical position of the handle */
}

.irs-single,
.irs-from,
.irs-to {
  font-size: 14px !important;
  font-weight: bold;
  top: -4px !important;   /* Move the value label up slightly */
  color: #3c2372;
}

.irs-min,
.irs-max {
  display: none !important;   /* Hide the min/max endpoint labels */
}

/* ---------------------------------------------------------
   Navbar (the top navigation bar)
   Controls height and vertical alignment of the nav elements.
   --------------------------------------------------------- */

.navbar {
  min-height: 50px !important;
  height: 50px !important;
  padding-top: 8px !important;
  padding-bottom: 8px !important;
}

/* Make nav items vertically centred inside the navbar */
.navbar-brand,
.navbar-nav,
.navbar-nav > li,
.navbar-header {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
}

.navbar-brand {
  line-height: 20px !important;
  padding: 0 !important;
  margin: 0 !important;
}

.navbar-nav {
  margin: 0 !important;
}

/* Style individual nav links — white text, no border, flex-aligned */
.navbar-nav > li > a {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
  line-height: 20px !important;
  padding: 7px 12px !important;
  background: none !important;
  border: none !important;
  color: white !important;
  font-weight: normal !important;
  text-decoration: none !important;
}

.navbar-nav > li > a:hover {
  color: rgba(255, 255, 255, 0.8) !important;  /* Slightly dimmed on hover */
}

/* Underline the currently active nav page */
.navbar-nav > li.active > a,
.navbar-nav > li > a.active,
.navbar-nav > li[class*="active"] > a {
  text-decoration: underline !important;
}

.navbar-nav > li.active > a:hover,
.navbar-nav > li > a.active:hover,
.navbar-nav > li[class*="active"] > a:hover {
  text-decoration: underline !important;
}

/* ---------------------------------------------------------
   Layout Spacing
   Reduces default padding/margins so the dashboard content
   fills more of the available screen space.
   --------------------------------------------------------- */

body {
  padding-top: 10px !important;   /* Small gap below the navbar */
}

.container-fluid {
  padding-top: 2px !important;
  padding-left: 15px !important;
  padding-right: 15px !important;
  margin: 0 !important;
  width: 100% !important;   /* Use full browser width */
}

.row {
  margin-left: 0 !important;
  margin-right: 0 !important;
  width: 100% !important;
}

/* Tighten column gutters */
.col-md-4,
.col-md-8 {
  padding-left: 5px !important;
  padding-right: 5px !important;
}

/* Tighten bottom spacing on chart panels */
.section.level3,
.chart-wrapper,
.chart-stage,
.html-widget {
  margin-bottom: 5px !important;
  padding-bottom: 5px !important;
}

/* ---------------------------------------------------------
   Crosstalk Checkbox Labels
   Makes the category labels next to checkbox filters slightly 
   larger for readability.
   --------------------------------------------------------- */

.crosstalk-input label {
  font-size: 14.5px !important;
}

</style>


```{r setup, include=FALSE}
# SETUP CHUNK
# 'include=FALSE' means this chunk runs silently — its code and
# output are NOT shown in the final dashboard.
#
# Here we:
#   1. Load all required R packages
#   2. Set a filename prefix used throughout the script
#   3. Apply a global option to suppress scientific notation
#   4. Apply a CSS rule to allow long column header text to wrap

# Load packages
library(drc)           # Dose-response curve fitting (4-parameter logistic models) - this package must be loaded first
library(data.table)    # Fast data manipulation (used here mainly for compatibility)
library(DT)            # Creates interactive HTML data tables (DataTables.js)
library(patchwork)     # Combines multiple ggplot2 plots into one figure
library(plotly)        # Creates interactive, web-based plots
library(tidyverse)     # Core data science packages: dplyr, ggplot2, readr, tidyr, etc.
library(flexdashboard) # Turns R Markdown into an interactive web dashboard
library(crosstalk)     # Enables linked filtering across multiple HTML widgets
library(htmltools)     # Tools for building HTML elements from R

# Set filename prefix
prefix <- "VCFG_Data02"

# Suppress scientific notation
# This forces use of decimals to improve readability
options(scipen = 999)

# Allow DataTable column headers to wrap onto multiple lines so full names are visible
tags$style(HTML("
  table.dataTable thead th {
    white-space: normal !important;
    word-break: break-word;
  }
"))
```

```{r input per well data}
# ============================================================
# DATA IMPORT
# Read in the two main data files from the 'deidentified' data folder.
# 
# - 'data_analysed_well':  One row per well (i.e. per compound × dose × cell line).
#                          Contains raw and processed measurements at each dose.
# - 'data_analysed_comp':  One row per compound × cell line combination.
#                          Contains curve-fit summary statistics (e.g. IC50, AUC, GR50).
# ============================================================

data_analysed_well <- read_csv(paste0("../data/VCFG_Data02/", prefix, "_analysed_data_well.csv"))  

data_analysed_comp <- read_csv(paste0("../data/VCFG_Data02/", prefix, "_analysed_data_compound.csv")) 
```

```{r, prep data}
# ============================================================
# DATA PREPARATION
# Clean and restructure the imported data ready for display 
# and filtering in the dashboard.
# ============================================================

# --- Define ordered factor levels ---
# Factors with a defined order ensure that categories appear in a 
# meaningful sequence in filters and plots, rather than alphabetically.

# Viability bins 
vbin_order <- c("High", "Normal", "Moderate", "Low", "Very Low")

# Growth inhibition bins 
gibin_order <- c("Cytotoxic", "Strong Inhibition", "Partial Inhibition", "No Effect", "Growth Stimulation")

# Death type bins
dtbin_order <- c("No Effect", "Partially Cytostatic", "Cytotoxic")

# Prepare the well-level (dose-level) data
data_all_well <- data_analysed_well %>%
  arrange(Compound_Name, Cell_Line, desc(Concentration)) %>% 
  mutate(
    Concentration = paste0(Concentration, units),
    Viability_Bin = factor(Viability_Bin, levels = vbin_order), # converting columns to correct order as defined above
    Growth_Inhibition_Bin = factor(Growth_Inhibition_Bin, levels = gibin_order),
    # Create a unique identifier for each well by pasting together key columns.
    # Used internally for linking widgets (not displayed to users).
    Unique_ID = paste(Cell_Line, Compound_Name, Concentration, sep = "_"),
    # Eval_Key is the crosstalk linking key: rows with the same Eval_Key 
    # are treated as belonging to the same compound group when filtering.
    Eval_Key = Compound_Name
  )

# Prepare the compound-level (curve-fit summary) data
data_all_comp <- data_analysed_comp %>%
  arrange(Compound_Name, Cell_Line) %>% 
  mutate(
    GRmax_Death_Type = factor(GRmax_Death_Type, levels = dtbin_order),
    GRinf_Death_Type = factor(GRinf_Death_Type, levels = dtbin_order),
    # Unique ID for each compound × cell line combination
    Unique_ID = paste(Cell_Line, Compound_Name, sep = "_"),
    # Linking key for crosstalk filtering (see below)
    Eval_Key = Compound_Name
  ) %>% 
  # Keep only the columns needed for display, in the desired column order.
  # This removes any extra columns that came in with the raw CSV.
  select(Cell_Line, Compound_Name, concentration_points, ctrl_cell_doublings,
         fit_rel_cell, AUC, Emax, Einf, IC50, EC50, h, r2_rel_cell, pval_rel_cell, flat_fit_rel_cell, 
         fit_GR, GR_AOC, GRmax, GRmax_Death_Type, GRinf, GRinf_Death_Type, GR50, GEC50, h_GR, r2_GR, pval_GR, flat_fit_GR, 
         Unique_ID, Eval_Key)
```

```{r shared data, include=TRUE}
# ============================================================
# CROSSTALK SHARED DATA OBJECTS
#
# Crosstalk allows multiple HTML widgets (tables, sliders, checkboxes)
# to communicate with each other — when a filter is applied, all 
# widgets sharing the same group update simultaneously.
#
# SharedData$new() wraps a dataframe into a crosstalk-aware object.
# Key arguments:
#   group = "drug1"  : All widgets in the same group are linked together.
#                      Filtering one widget filters all others in the group.
#   key = ~Eval_Key  : The column used to match rows across datasets.
#                      Rows with the same key value are highlighted/filtered together.
#                      Here, Eval_Key = Compound_Name, so selecting a compound
#                      highlights it across all linked widgets.
#
# We create separate shared data objects for each cell line so that the
# Cell Line X sliders only filter Cell Line X data, and Cell Line Y 
# sliders only filter Cell Line Y data — while still being in the same
# group so that compound-level selections stay in sync.
# ============================================================

# Cell line-specific shared data objects (for the sidebar filters)

# Well-level (dose-level) data, split by cell line
data_cellx_shared_well <- data_all_well %>%
  filter(Cell_Line == "Cell_Line_X") %>%   # Keep only Cell Line X rows
  SharedData$new(., group = "drug1", key = ~Eval_Key)

data_celly_shared_well <- data_all_well %>%
  filter(Cell_Line == "Cell_Line_Y") %>%   # Keep only Cell Line Y rows
  SharedData$new(., group = "drug1", key = ~Eval_Key)

# Compound-level (curve-fit) data, split by cell line
data_cellx_shared_comp <- data_all_comp %>%
  filter(Cell_Line == "Cell_Line_X") %>% 
  SharedData$new(., group = "drug1", key = ~Eval_Key)

data_celly_shared_comp <- data_all_comp %>%
  filter(Cell_Line == "Cell_Line_Y") %>%  
  SharedData$new(., group = "drug1", key = ~Eval_Key)

# Full (all cell lines) shared data objects (for the downloadable data tables)
# These include both cell lines and are used to populate the main data tables.
# They share the same group so table rows update when sidebar filters are applied.

data_all_shared_well <- data_all_well %>%
  SharedData$new(., group = "drug1", key = ~Eval_Key)

data_all_shared_comp <- data_all_comp %>%
  SharedData$new(., group = "drug1", key = ~Eval_Key)
```

<!-- ============================================================
  DASHBOARD LAYOUT
  In flexdashboard, the layout is controlled by Markdown headings.
  
  '## Column' starts a new column panel.
  'data-width=275' sets the pixel width of the column. This can be adjusted if more/fewer columns are desired.
  '### Panel Title' creates a named panel within a column.
  ============================================================ -->

Column {data-width=275}
-----------------------------------------------------------------------

### Hit Identification Filters

<!-- Static Markdown text displayed at the top of the sidebar.
     Explains to the user how the filters work. -->

**Use the filters below to refine your hit list:**

* Slider bars: Move the handles on the sliders bars to select compounds with values within the desired range. The **purple** area of the slider bar indicates the value range you would like to keep.
* Tick boxes: Tick to select compounds in specific categories. The **ticked** boxes indicate the categories you would like to keep.


***Endpoint Relative Cell Count:***

**Cell Line X**
```{r}
# ============================================================
# SIDEBAR FILTERS — AUC Sliders
#
# filter_slider() creates an interactive range slider. When the 
# user moves the handles, only compounds with AUC values within 
# the selected range remain visible in all linked widgets.
#
# Arguments:
#   id          : A unique HTML ID for this filter element
#   label       : Text label displayed above the slider
#   sharedData  : The SharedData object this filter acts on
#   ~AUC        : The column to filter 
#   step        : The minimum increment the slider moves by
# ============================================================
filter_slider(
  id = "filter1",
  label = "AUC (low values = toxicity):",
  sharedData = data_cellx_shared_comp,   # Acts on Cell Line X compound data
  ~AUC,
  step = 0.01
)
```

**Cell Line Y**
```{r}
# Same AUC slider as above, but linked to Cell Line Y compound data
filter_slider(
  id = "filter2",
  label = "AUC (low values = toxicity):",
  sharedData = data_celly_shared_comp,   # Acts on Cell Line Y compound data
  ~AUC,
  step = 0.01
)
```

***Growth Rate-Adjusted Inhibition:***

**Cell Line X**
```{r}
# ============================================================
# SIDEBAR FILTERS — GRmax Checkboxes (Cell Line X)
#
# filter_checkbox() creates a set of tick boxes, one per category.
# When a box is ticked, only compounds in that category are shown.
# Unticking all boxes shows no compounds (not all compounds).
#
# Arguments:
#   allLevels = TRUE : Show all factor levels, even if none currently 
#                      match (ensures the full list is always visible)
#   inline = TRUE    : Place checkboxes side by side rather than stacked
#   columns = 1      : (with inline=TRUE) Arrange into 1 column
# ============================================================
filter_checkbox(
  id = "filter3",
  label = "GRmax Death Type:",
  sharedData = data_cellx_shared_comp,
  ~GRmax_Death_Type,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)

# AUC slider for growth rate data (Cell Line X)
filter_slider(
  id = "filter4",
  label = "GR AOC (high values = toxicity):",
  sharedData = data_cellx_shared_comp,
  ~GR_AOC,
  step = 0.01
)
```

**Cell Line Y**
```{r}
# Same checkbox and slider filters as above, linked to Cell Line Y data
filter_checkbox(
  id = "filter5",
  label = "GRmax Death Type:",
  sharedData = data_celly_shared_comp,
  ~GRmax_Death_Type,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)

filter_slider(
  id = "filter6",
  label = "GR AOC (high values = toxicity):",
  sharedData = data_celly_shared_comp,
  ~GR_AOC,
  step = 0.01
)
```

```{r}
# ============================================================
# RESET BUTTON
# A simple HTML button that reloads the entire page when clicked,
# which clears all filter selections back to their defaults.
# tags$button() is from htmltools — it generates a raw HTML <button>.
# ============================================================
tags$button(
  "Reset All Filters",
  class = "btn btn-default",       # Bootstrap button styling class
  onclick = "window.location.reload();",  # JavaScript: reload page on click
  style = "margin: 10px 0;"
)
```

<!-- Note the increased width for this column that contains tables -->

Column {data-width=660} 
-----------------------------------------------------------------------

### Compound-Level Data Table

```{r}
# ============================================================
# COMPOUND-LEVEL DATA TABLE
#
# datatable() from the DT package creates an interactive HTML table
# with sorting, searching, pagination, and export buttons.
#
# Key arguments:
#   data            : The SharedData object to display (linked to filters)
#   extensions      : Extra DataTables plugins to enable
#     'Buttons'       : Adds Copy/Excel/column toggle buttons above the table
#     'FixedColumns'  : Freezes the leftmost columns during horizontal scrolling
#     'ColReorder'    : Lets the user drag-and-drop columns to reorder them
#   rownames = FALSE : Don't show row numbers in a separate first column
#   class            : CSS class applied to the table element
# ============================================================
datatable(
  data_all_shared_comp, 
  extensions = c('Buttons', 'FixedColumns', 'ColReorder'),
  rownames = FALSE,
  escape = FALSE,       # Allow HTML formatting in cell content (not used here but safe to include)
  class = 'cell-border',
  options = list(
    
    # --- DOM layout string ---
    # Controls which DataTable UI elements appear and in what order.
    # B = Buttons, s = search box, f = filter, r = processing indicator, 
    # t = table, i = info text (showing X of Y entries)
    dom = 'Bsfrti',
    
    # --- drawCallback: JavaScript run each time the table redraws ---
    # This custom JS adds a thick purple border above the first row of 
    # each new compound group, making it visually clear when the compound 
    # changes in the table (since each compound has 2 rows: one per cell line).
    drawCallback = JS(
      "function(settings) {",
      "  var api = this.api();",
      "",
      # Remove any borders added in the previous draw (prevents duplicates on redraw)
      "  api.rows({ page: 'current' }).nodes().each(function(row) {",
      "    $('td', row).css('border-top', '');",
      "  });",
      "",
      # Get the rows and data currently visible on this page
      "  var rows = api.rows({ page: 'current' }).nodes();",
      "  var data = api.rows({ page: 'current' }).data();",
      "",
      # Find the index of the Compound_Name column (needed for comparison below)
      "  var compoundIdx = api.column('Compound_Name:name').index();",
      "",
      # Loop through each row; if the compound name differs from the row above,
      # draw a thick purple top border — marking the start of a new compound
      "  for (var i = 0; i < data.length; i++) {",
      "    if (i === 0 || data[i][compoundIdx] !== data[i - 1][compoundIdx]) {",
      "      $('td', rows[i]).css('border-top', '2px solid #3c2372');",
      "    }",
      "  }",
      "}"
    ),
    
    # --- columnDefs: Controls the behaviour of specific columns ---
    columnDefs = list(
      # Centre-align all column content
      list(className = 'dt-center', targets = "_all"),
      
      # Give the Compound_Name column a named reference for the JS above
      # which(colnames(...) == "Compound_Name") finds the zero-based column index
      list(name = 'Compound_Name', targets = which(colnames(data_all_shared_comp) == "Compound_Name")),
      
      # Hide detailed/technical columns by default — users can show them 
      # via the 'Show/Hide Columns' button if needed.
      list(visible = FALSE, targets = c('concentration_points',
                                        'ctrl_cell_doublings',
                                        'fit_GR',
                                        'flat_fit_GR',
                                        'GR50',
                                        'GRmax',
                                        'GEC50',
                                        'GRinf',
                                        'GRinf_Death_Type',
                                        'h_GR',
                                        'r2_GR',
                                        'pval_GR',
                                        'fit_rel_cell',
                                        'flat_fit_rel_cell',
                                        'Emax',
                                        'Einf',
                                        'IC50',
                                        'EC50',
                                        'h',
                                        'r2_rel_cell',
                                        'pval_rel_cell'
                                        )),
      # Hide the linking columns completely — they're used for crosstalk 
      # filtering only and should never appear in the UI.
      # 'noVis' class also excludes them from the Show/Hide Columns menu.
      list(visible = FALSE, targets = c('Unique_ID', 'Eval_Key'), className = 'noVis')),
    
    pageLength = 20,   # Show 20 rows per page by default
    
    # --- Export/utility buttons ---
    buttons = list(
      # Copy visible table data to the clipboard as tab-separated text
      list(extend = 'copyHtml5',
           text = "Copy to Clipboard"),
      
      # Export visible table data to an Excel (.xlsx) file
      # filename: the default name for the downloaded file
      # title = NULL: don't add a title row in the Excel file
      list(extend = 'excel',
           filename = paste0(prefix, "_Selected_Compounds_Compound_Level_Data"), 
           title = NULL),
      
      # A dropdown menu to show/hide individual columns
      # columns = ':not(.noVis)' excludes the hidden crosstalk columns from the menu
      list(extend = 'colvis',
           text = 'Show/Hide Columns',
           columns = ':not(.noVis)')),
    
    searching = TRUE,        # Show a text search box above the table
    scrollX = TRUE,          # Enable horizontal scrolling for wide tables
    scrollCollapse = TRUE,   # Don't add empty space below a short table
    autoWidth = FALSE,       # Don't auto-resize columns (use fixed widths)
    paging = TRUE,           # Break long tables into pages
    colReorder = TRUE,       # Allow users to reorder columns by dragging
    
    # Freeze the first 3 columns so they stay fixed in place when scrolling right.
    fixedColumns = list(leftColumns = 3)),
  
  elementId = "data_download_table_2"   # A unique HTML ID for this table element
) %>% 

  # --- Conditional cell background colours ---

  # Colour the Cell_Line column: blue for Cell Line X, red/pink for Cell Line Y
  formatStyle(
    'Cell_Line',
    backgroundColor = styleEqual(
      c('Cell_Line_X', 'Cell_Line_Y'),   
      c("#A8D4EE",                         
        "#E08C86") 
    )
  ) %>% 
  
  # Colour the GRmax_Death_Type column using a traffic-light scheme:
  # red = cytotoxic (most effective), yellow = partial, green = no effect
  formatStyle(
    columns = 'GRmax_Death_Type',
    backgroundColor = styleEqual(
      c('Cytotoxic', 'Partially Cytostatic', 'No Effect'),
      c('#ffcdd2',      # Soft red
        '#fff9c4',      # Soft yellow
        '#f1f8e9')      # Soft green
    )
  ) 
```

### Dose-Level Data Table

```{r}
# ============================================================
# DOSE-LEVEL DATA TABLE
#
# Same structure as the compound-level table above, but showing 
# one row per well (compound × dose × cell line) instead of 
# one row per compound × cell line.
#
# Key differences:
#   - Uses data_all_shared_well (well-level data)
#   - Different columns hidden/shown
#   - Additional colour-coding for Concentration, Viability_Bin,
#     and Growth_Inhibition_Bin columns
# ============================================================
datatable(
  data_all_shared_well, 
  extensions = c('Buttons', 'FixedColumns', 'ColReorder'),
  rownames = FALSE,
  escape = FALSE,
  class = 'cell-border',
  options = list(
    dom = 'Bsfrti',
    
    # Same compound-grouping border logic as the compound table above
    drawCallback = JS(
      "function(settings) {",
      "  var api = this.api();",
      "",
      "  api.rows({ page: 'current' }).nodes().each(function(row) {",
      "    $('td', row).css('border-top', '');",
      "  });",
      "",
      "  var rows = api.rows({ page: 'current' }).nodes();",
      "  var data = api.rows({ page: 'current' }).data();",
      "",
      "  var compoundIdx = api.column('Compound_Name:name').index();",
      "",
      "  for (var i = 0; i < data.length; i++) {",
      "    if (i === 0 || data[i][compoundIdx] !== data[i - 1][compoundIdx]) {",
      "      $('td', rows[i]).css('border-top', '2px solid #3c2372');",
      "    }",
      "  }",
      "}"
    ),
    columnDefs = list(
      list(className = 'dt-center', targets = "_all"),
      list(name = 'Compound_Name', targets = which(colnames(data_all_shared_well) == "Compound_Name")),
      # Hide raw/internal columns that aren't needed for the main view
      list(visible = FALSE, targets = c('log10_concentration',
                                        'units',
                                        'time',
                                        'cell_count__time0', 
                                        'cell_count__endpoint_ctrl',
                                        'ctrl_cell_doublings',
                                        'cell_count__endpoint'
                                        )),
      list(visible = FALSE, targets = c('Unique_ID', 'Eval_Key'), className = 'noVis')),
    pageLength = 20,
    buttons = list(
      list(extend = 'copyHtml5',
           text = "Copy to Clipboard"),
      list(extend = 'excel',
           filename = paste0(prefix, "_Selected_Compounds_Dose_Level_Data"), 
           title = NULL),
      list(extend = 'colvis',
           text = 'Show/Hide Columns',
           columns = ':not(.noVis)')),
    searching = TRUE,
    scrollX = TRUE,
    scrollCollapse = TRUE,
    autoWidth = FALSE,
    paging = TRUE,
    colReorder = TRUE,
    fixedColumns = list(leftColumns = 3)),
  elementId = "data_download_table_1") %>% 
  
  # --- Alternating grey shading for the Concentration column ---
  # Each dose level gets a slightly different shade of grey so rows 
  # within the same compound are visually grouped by dose.
  # styleEqual maps specific Concentration values to specific colours.
  # valueColumns = 'Concentration' means the colour is driven by the 
  # value in Concentration, even though the columns styled are Compound_Name 
  # and Concentration together.
  formatStyle(
    c('Compound_Name', 'Concentration'),
    backgroundColor = styleEqual(
      c("0.5uM", "0.15uM", "0.047uM", "0.015uM"),
      c("#e9e9e9", "#f0f0f0", "#f7f7f7", "#ffffff")
    ),
    valueColumns = 'Concentration'
  ) %>%
  
  # Colour the Cell_Line column (same scheme as compound table)
  formatStyle(
    'Cell_Line',
    backgroundColor = styleEqual(
      c('Cell_Line_X', 'Cell_Line_Y'),
      c("#A8D4EE", 
        "#E08C86") 
    )
  ) %>%
  
  # Traffic-light colours for the Viability_Bin column
  formatStyle(
    columns = 'Viability_Bin',
    backgroundColor = styleEqual(
      c('Very Low', 'Low', 'Moderate', 'Normal', 'High'),
      c('#ffcdd2',    # Red (very low viability — most toxic)
        '#ffccbc',    # Orange
        '#fff9c4',    # Yellow
        '#f1f8e9',    # Light green
        '#e1f5fe')    # Light blue (high viability — healthy)
    )
  ) %>% 
  
  # Traffic-light colours for the Growth_Inhibition_Bin column
  formatStyle(
    columns = 'Growth_Inhibition_Bin',
    backgroundColor = styleEqual(
      c('Cytotoxic', 'Strong Inhibition', 'Partial Inhibition', 'No Effect', 'Growth Stimulation'),
      c('#ffcdd2',    # Red (cytotoxic)
        '#ffccbc',    # Orange
        '#fff9c4',    # Yellow
        '#f1f8e9',    # Light green
        '#e1f5fe')    # Light blue (growth stimulation)
    )
  )
```

Column
-----------------------------------------------------------------------

### Single Compound View: Dose-Response Plot

```{r}
# ============================================================
# DOSE-RESPONSE CURVE FITTING
#
# For the "Single Compound View" panel, we need smooth dose-response
# curves to display alongside the raw data points. This chunk:
#
#   1. Prepares the well-level data for plotting
#   2. Fits 4-parameter logistic (4PL) models to each compound × cell line
#   3. Generates predictions from those models over a fine concentration grid
#
# The 4PL model (L.4 in the drc package) describes a sigmoidal curve 
# with four parameters: slope, lower asymptote, upper asymptote, and ED50.
# ============================================================

# Prepare data: convert concentrations back to numeric and add a Unique_ID
data_single_plot <- data_analysed_well %>%
  mutate(Concentration = as.numeric(Concentration),
         log10_concentration = as.numeric(log10_concentration),
         Unique_ID = paste(Cell_Line, Compound_Name, sep = "_"))

# --- Build an x-axis grid for plotting smooth curves ---
# seq() creates 200 evenly-spaced points between the min and max 
# log10(concentration) values in the data. Using log10 scale gives 
# equal visual spacing across orders of magnitude.
log_grid <- seq(min(data_single_plot$log10_concentration, na.rm = TRUE),
                max(data_single_plot$log10_concentration, na.rm = TRUE),
                length.out = 200)

# --- Fit relative cell count curves ---
# For each compound × cell line combination, fit a 4PL model and 
# generate predicted values across the log_grid.
preds_rel_cell <- data_single_plot %>%
  group_by(Compound_Name, Cell_Line) %>%
  group_modify(~ {
    
    # Fit the 4PL dose-response model.
    # drm() = dose-response model; L.4() = 4-parameter log-logistic function
    # tryCatch() prevents the whole script crashing if a model fails to fit —
    # it catches the error and returns NULL instead.
    model <- tryCatch(
      drm(
        rel_cell_count ~ Concentration,   # Formula: y ~ x
        data = .x,                         # .x is the current group's data
        fct = L.4(names = c("slope", "lower", "upper", "ED50"))),
      error = function(e) NULL)            # If fitting fails, return NULL
    
    # If the model failed, return an empty tibble (no predictions for this group)
    if (is.null(model)) return(tibble())
    
    # Otherwise, predict rel_cell_count at each point in log_grid.
    # 10^log_grid converts back from log10 scale to actual concentration 
    # values, which the model expects.
    tibble(
      log10_concentration = log_grid,
      rel_cell_count = predict(
        model,
        newdata = data.frame(
          Concentration = 10^log_grid)
      )
    )
  }) %>%
  ungroup()

# --- Fit GR value curves (growth rate-adjusted inhibition) ---
# Identical process as above, but fitting GRvalue instead of rel_cell_count
preds_gr <- data_single_plot %>%
  group_by(Compound_Name, Cell_Line) %>%
  group_modify(~ {
    model <- tryCatch(
      drm(
        GRvalue ~ Concentration,
        data = .x,
        fct = L.4(names = c("slope", "lower", "upper", "ED50"))),
      error = function(e) NULL)
    if (is.null(model)) return(tibble())
    
    tibble(
      log10_concentration = log_grid,
      GRvalue = predict(
        model,
        newdata = data.frame(
          Concentration = 10^log_grid)
      )
    )
  }) %>%
  ungroup()
```

```{r shared single comp data, include=TRUE}
# ============================================================
# SHARED DATA FOR THE SINGLE-COMPOUND PLOT
#
# A separate crosstalk group ("drug2") is used here, independent 
# from the table filters ("drug1") - this keeps the filtering separate.
#
# The key is ~Compound_Name (not ~Eval_Key as above) because the 
# filter_select() dropdown below links directly to compound names,
# and we only need to display one compound at a time in the plot.
#
# Three linked datasets are needed:
#   - Raw data points (plot_data_single_shared)
#   - Predicted curve for relative cell count (preds_rel_cell_shared)
#   - Predicted curve for GR value (preds_gr_shared)
# All three update together when the user selects a different compound.
# Therefore all three need to be made into SharedData objects with 
# the same group and key.
# ============================================================

plot_data_single_shared <- data_single_plot %>%
  SharedData$new(., group = "drug2", key = ~Compound_Name)

preds_rel_cell_shared <- preds_rel_cell %>%
  SharedData$new(., group = "drug2", key = ~Compound_Name)

preds_gr_shared <- preds_gr %>%
  SharedData$new(., group = "drug2", key = ~Compound_Name)
```

```{r}
# ============================================================
# COMPOUND SELECTOR DROPDOWN
#
# filter_select() creates a searchable dropdown menu.
# When the user selects a compound, the "drug2" group updates,
# and only data for that compound is shown in the plots below.
#
# multiple = FALSE: only one compound can be selected at a time
# ============================================================
filter_select(
  id = "filter7",
  label = "Select a compound to plot:",
  sharedData = plot_data_single_shared,
  ~Compound_Name,
  allLevels = FALSE,
  multiple = FALSE
)
```

```{js}
// ============================================================
// JAVASCRIPT: Set default compound on page load
//
// By default, filter_select() shows no selection (all compounds visible).
// This JS snippet runs after the page loads and pre-selects a specific
// compound ("Test_Compound01") so the plots display immediately on open.
//
// It also removes the blank "All" option from the dropdown menu so 
// the user must always have a compound selected.
//
// setTimeout(..., 100) introduces a 100ms delay to ensure the 
// selectize widget has fully initialised before we try to modify it.
// ============================================================
$(document).ready(function() {
  setTimeout(function() {
    var selectElement = $("#filter7 .selectized").first();
    if (selectElement.length) {
      var selectize = selectElement[0].selectize;
      
      // Pre-select the first compound
      selectize.setValue("Test_Compound01", false);
      
      // Remove the empty "" option (the "All" / blank option)
      selectize.removeOption("");
    }
    
  }, 100);
});
```

```{r}
# ============================================================
# DOSE-RESPONSE PLOTS (Plotly)
#
# Two stacked interactive plots are created here:
#   1. p_rel: Endpoint Relative Cell Count vs Concentration
#   2. p_gr:  Growth Rate-Adjusted Inhibition (GR Value) vs Concentration
#
# Both plots are then combined into a single stacked layout using subplot().
# ============================================================

# --- Colour palette ---
# Named vector mapping cell line names to specific hex colours.
# Using a named vector ensures the colours are consistently assigned 
# regardless of the order data appears.
condition_colors <- c("Cell_Line_X" = "#1F78B4",   # Blue
                      "Cell_Line_Y"  = "#B22222")   # Dark red

# --- X-axis tick marks ---
# Pull the unique concentration values from the data to use as tick positions.
# We plot on a log10 scale (x = log10_concentration) but label ticks with
# the original concentration values (e.g. 0.015, 0.047, 0.15, 0.5) for clarity.
x_breaks <- data_single_plot %>%
  distinct(Concentration) %>%
  arrange(Concentration) %>%
  pull(Concentration)

x_tickvals <- log10(x_breaks)         # Tick positions on the log10 scale
x_ticktext <- format(x_breaks, trim = TRUE)  # Formatted labels (no leading space)

# ============================================================
# PLOT 1: Endpoint Relative Cell Count
# ============================================================

p_rel <- plot_ly() %>%

  # --- Raw data points ---
  # add_markers() adds scatter points. Each cell line gets its own colour.
  # 'legendgroup' links points and lines for the same cell line so clicking 
  # the legend toggles both simultaneously.
  add_markers(
    data = plot_data_single_shared,          # Crosstalk shared data (filters by compound)
    x = ~log10_concentration,
    y = ~rel_cell_count,
    color = ~Cell_Line,                      # Colour by cell line
    colors = condition_colors,
    marker = list(size = 8),
    name = ~Cell_Line,                       # Legend entry label
    legendgroup = ~Cell_Line,                # Group with the line for the same cell line
    showlegend = TRUE,                       # Show this series in the legend
    hovertemplate = ~paste0(                 # Custom hover tooltip text
      "<b>", Compound_Name, "</b>",
      "<br>Cell Line: ", Cell_Line,
      "<br>Conc: ", Concentration, " µM",
      "<br>Relative cell count: ", round(rel_cell_count, 2),
      "<extra></extra>"                      # Removes the default plotly trace name box
    )
  ) %>% 

  # --- Fitted dose-response curves ---
  # add_lines() draws the smooth predicted curves from the 4PL model.
  # showlegend = FALSE prevents duplicate legend entries (already shown by markers).
  add_lines(
    data = preds_rel_cell_shared,
    x = ~log10_concentration,
    y = ~rel_cell_count,
    color = ~Cell_Line,
    colors = condition_colors,
    line = list(width = 4),                  # Thicker line than default
    name = ~Cell_Line,
    legendgroup = ~Cell_Line,                # Same group as the markers above
    showlegend = FALSE,
    hovertemplate = ~paste0(
      "<b>", Cell_Line, "</b>",
      "<extra></extra>"
    )
  ) %>%

  # --- Axis and layout formatting ---
  layout(
    xaxis = list(
      title = list(text = "Compound Concentration (µM)", font = list(size = 16)),
      tickvals = x_tickvals,     # Where to place tick marks
      ticktext = x_ticktext,     # What label to show at each tick
      tickfont = list(size = 14),
      zeroline = FALSE           # Don't draw a vertical line at x = 0
    ),
    yaxis = list(
      title = list(text = "Cell Count Relative to DMSO", font = list(size = 16)),
      range = c(0, 1.75),        # Fix y-axis from 0 to 1.75 
      tickvals = seq(0, 1.75, by = 0.25),
      tickfont = list(size = 14),
      zeroline = FALSE
    ),
    legend = list(font = list(size = 14))
  )

# --- Reference lines ---
# These are added separately as layout 'shapes' (drawn as fixed lines, 
# not as data traces) so they don't appear in the legend.
x_min <- min(data_single_plot$log10_concentration, na.rm = TRUE)
x_max <- max(data_single_plot$log10_concentration, na.rm = TRUE)

p_rel <- p_rel %>%
  layout(
    shapes = list(
      # Dashed horizontal line at y = 1 (the "no effect" reference — 
      # i.e. same cell count as DMSO control)
      list(
        type = "line",
        x0 = x_min, x1 = x_max,
        y0 = 1, y1 = 1,
        line = list(dash = "dot", color = "grey")
      ),
      # Solid vertical line at the leftmost x position (acts as y-axis border)
      list(
        type = "line",
        x0 = x_min, x1 = x_min,
        y0 = 0, y1 = 1.75,
        line = list(color = "black", width = 1)
      ),
      # Solid horizontal line at y = 0 (acts as x-axis border)
      list(
        type = "line",
        x0 = x_min, x1 = x_max,
        y0 = 0, y1 = 0,
        line = list(color = "black", width = 1)
      )
    )
  )

# ============================================================
# PLOT 2: Growth Rate-Adjusted Inhibition (GR Value)
# Identical structure to p_rel, with two differences:
#   - Uses GRvalue on the y-axis (range: -1.1 to 2)
#   - GR values below 0 indicate cytotoxicity; above 1 = growth stimulation
# ============================================================
p_gr <- plot_ly() %>%

  add_markers(
    data = plot_data_single_shared,
    x = ~log10_concentration,
    y = ~GRvalue,
    color = ~Cell_Line,
    colors = condition_colors,
    marker = list(size = 8),
    name = ~Cell_Line,
    legendgroup = ~Cell_Line,
    showlegend = TRUE,
    hovertemplate = ~paste0(
      "<b>", Compound_Name, "</b>",
      "<br>Cell Line: ", Cell_Line,
      "<br>Conc: ", Concentration, " µM",
      "<br>GR value: ", round(GRvalue, 2),
      "<extra></extra>"
    )
  ) %>% 

  add_lines(
    data = preds_gr_shared,
    x = ~log10_concentration,
    y = ~GRvalue,
    color = ~Cell_Line,
    colors = condition_colors,
    line = list(width = 4),
    name = ~Cell_Line,
    legendgroup = ~Cell_Line,
    showlegend = FALSE,
    hovertemplate = ~paste0(
      "<b>", Cell_Line, "</b>",
      "<extra></extra>"
    )
  ) %>%

  layout(
    xaxis = list(
      title = list(text = "Compound Concentration (µM)", font = list(size = 16)),
      tickvals = x_tickvals,
      ticktext = x_ticktext,
      tickfont = list(size = 14),
      zeroline = FALSE
    ),
    yaxis = list(
      title = list(text = "GR Value Relative to DMSO", font = list(size = 16)),
      range = c(-1.1, 2),          # Extended range: GR < -1 = extreme cytotoxicity
      tickvals = seq(-1, 2, by = 0.5),
      tickfont = list(size = 14),
      zeroline = FALSE
    ),
    legend = list(font = list(size = 14))
  )

p_gr <- p_gr %>%
  layout(
    shapes = list(
      list(
        type = "line",
        x0 = x_min, x1 = x_max,
        y0 = 1, y1 = 1,
        line = list(dash = "dot", color = "grey")
      ),
      list(
        type = "line",
        x0 = x_min, x1 = x_min,
        y0 = -1.1, y1 = 2,
        line = list(color = "black", width = 1)
      ),
      list(
        type = "line",
        x0 = x_min, x1 = x_max,
        y0 = -1.1, y1 = -1.1,
        line = list(color = "black", width = 1)
      )
    )
  )

# Hide the legend for p_gr — the legend is already shown in p_rel above,
# and since both share legendgroups, one legend controls both plots.
p_gr <- p_gr %>%
  plotly::style(showlegend = FALSE)
```

```{r}
# ============================================================
# COMBINE PLOTS WITH subplot()
#
# subplot() stacks the two plotly objects vertically (nrows = 2).
# Each plot keeps its own x and y axes (shareX/shareY = FALSE).
# margin = 0.1 adds spacing between the two plots.
#
# Annotations are then added as panel titles positioned using 
# relative paper coordinates (xref/yref = "paper" means 0-1 
# maps to the full width/height of the combined figure).
# ============================================================
combined_plot <- subplot(
  p_rel,            # Top plot: relative cell count
  p_gr,             # Bottom plot: GR value
  nrows = 2,
  shareX = FALSE,   # Each plot has its own independent x-axis
  shareY = FALSE,   # Each plot has its own independent y-axis
  titleX = TRUE,    # Show x-axis titles on each subplot
  titleY = TRUE,    # Show y-axis titles on each subplot
  margin = 0.1      # Gap between the two plots (fraction of total height)
) %>%
  layout(
    # Add bold text labels above each subplot to act as panel titles.
    # x and y are paper-relative coordinates (0 = left/bottom, 1 = right/top).
    annotations = list(
      list(
        text = "<b>Endpoint Relative Cell Count</b>",
        x = 0.225,
        y = 1.02,           # Slightly above the top plot
        xref = "paper",
        yref = "paper",
        xanchor = "center",
        yanchor = "bottom",
        showarrow = FALSE,
        font = list(size = 17)
      ),
      list(
        text = "<b>Growth Rate-Adjusted Inhibition</b>",
        x = 0.225,
        y = 0.44,           # Positioned just above the bottom plot
        xref = "paper",
        yref = "paper",
        xanchor = "center",
        yanchor = "bottom",
        showarrow = FALSE,
        font = list(size = 17)
      )
    ),
    # Overall figure margins (pixels)
    margin = list(
      l = 80,     # Left margin (space for y-axis labels)
      r = 50,     # Right margin
      t = 50,     # Top margin
      b = 160     # Bottom margin (extra space for x-axis label)
    )
  )

# Render the combined plot in the dashboard panel
combined_plot
```
