---
# ============================================================
# YAML HEADER
# This block controls the overall appearance and layout of the 
# dashboard. It is not R code — it is read by the flexdashboard 
# package when knitting the document.
# ============================================================

title: "VCFG Dataset 1: Large-Scale Primary Screen"   # Title shown in the navbar
output: 
  flexdashboard::flex_dashboard:
    theme:                    # Custom colour theme using Bootstrap Sass variables - if using a preseet Bootswatch theme, specify it here
      bg: "white"             # Background colour of the dashboard
      fg: "#3c2372"           # Foreground (text) colour — a dark purple
      primary: "#3c2372"      # Primary accent colour (buttons, active elements)
      secondary: "#3c2372"    # Secondary accent colour
      success: "#3c2372"      # Colour for "success" state elements
    orientation: rows         # Panels are arranged in rows
    vertical_layout: fill     # Panels expand to fill the full height of the browser
---

<!-- ============================================================
  CSS STYLING BLOCK
  Raw HTML/CSS injected directly into the dashboard. This customises
  the appearance of specific elements beyond what the YAML theme covers.
  Each section targets a different part of the UI. Sections can be added
  or removed depending on your customisation requirements.
  ============================================================ -->

<style type="text/css">

/* ---------------------------------------------------------
   DataTables Buttons (Copy, Excel, Show/Hide Columns)
   Makes the row of export buttons compact and left-aligned,
   rather than stretching to fill the available width.
   --------------------------------------------------------- */

.dataTables_wrapper .dt-buttons {
  display: inline-flex !important;  /* Lay buttons out in a horizontal row */
  align-items: center;
  gap: 6px;                         /* Small gap between each button */
  width: auto !important;           /* Don't stretch to full container width */
  flex: 0 0 auto !important;
}

.dt-buttons .dt-button {
  font-size: 13px !important;
  padding: 4px 8px !important;
  background: #f8f8f8 !important;   /* Light grey button background */
  border: 1px solid #bbb !important;
  border-radius: 4px !important;
}

.dt-buttons .dt-button:hover {
  background: #eaeaea !important;   /* Slightly darker on hover */
  border-color: #999 !important;
}

/* ---------------------------------------------------------
   Chart Panel Titles
   The text shown above each panel in the dashboard.
   --------------------------------------------------------- */

.chart-title {  
  font-size: 19px;
  font-weight: bold;
  color: #3c2372;   /* Match the dashboard's purple theme colour */
}

/* ---------------------------------------------------------
   ionRangeSlider (the interactive slider filter widgets)
   The slider is created by the crosstalk filter_slider() function.
   These rules customise its colours and hide default labels that
   would otherwise clutter the UI.
   --------------------------------------------------------- */

.irs-grid-text {
  color: #3c2372;
  display: none !important;   /* Hide the numeric grid labels beneath the slider */
}

.irs-bar {
  background-color: #3c2372; /* Purple fill for the selected range */
  border: 1px solid #3c2372;
}

.irs-slider {
  background: #3c2372 !important;   /* The draggable handle colour */
  border: 2px solid #3c2372;
  width: 10px !important;
  height: 10px !important;
  top: 25px !important;             /* Vertical position of the handle */
}

.irs-single,
.irs-from,
.irs-to {
  font-size: 14px !important;
  font-weight: bold;
  top: -4px !important;   /* Move the value label up slightly */
  color: #3c2372;
}

.irs-min,
.irs-max {
  display: none !important;   /* Hide the min/max endpoint labels */
}

/* ---------------------------------------------------------
   Navbar (the top navigation bar)
   Controls height and vertical alignment of the nav elements.
   --------------------------------------------------------- */

.navbar {
  min-height: 50px !important;
  height: 50px !important;
  padding-top: 8px !important;
  padding-bottom: 8px !important;
}

/* Make nav items vertically centred inside the navbar */
.navbar-brand,
.navbar-nav,
.navbar-nav > li,
.navbar-header {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
}

.navbar-brand {
  line-height: 20px !important;
  padding: 0 !important;
  margin: 0 !important;
}

.navbar-nav {
  margin: 0 !important;
}

/* Style individual nav links — white text, no border, flex-aligned */
.navbar-nav > li > a {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
  line-height: 20px !important;
  padding: 7px 12px !important;
  background: none !important;
  border: none !important;
  color: white !important;
  font-weight: normal !important;
  text-decoration: none !important;
}

.navbar-nav > li > a:hover {
  color: rgba(255, 255, 255, 0.8) !important;  /* Slightly dimmed on hover */
}

/* Underline the currently active nav page */
.navbar-nav > li.active > a,
.navbar-nav > li > a.active,
.navbar-nav > li[class*="active"] > a {
  text-decoration: underline !important;
}

.navbar-nav > li.active > a:hover,
.navbar-nav > li > a.active:hover,
.navbar-nav > li[class*="active"] > a:hover {
  text-decoration: underline !important;
}

/* ---------------------------------------------------------
   Layout Spacing
   Reduces default padding/margins so the dashboard content
   fills more of the available screen space.
   --------------------------------------------------------- */

body {
  padding-top: 10px !important;   /* Small gap below the navbar */
}

.container-fluid {
  padding-top: 2px !important;
  padding-left: 15px !important;
  padding-right: 15px !important;
  margin: 0 !important;
  width: 100% !important;   /* Use full browser width */
}

.row {
  margin-left: 0 !important;
  margin-right: 0 !important;
  width: 100% !important;
}

/* Tighten column gutters */
.col-md-4,
.col-md-8 {
  padding-left: 5px !important;
  padding-right: 5px !important;
}

/* Tighten bottom spacing on chart panels */
.section.level3,
.chart-wrapper,
.chart-stage,
.html-widget {
  margin-bottom: 5px !important;
  padding-bottom: 5px !important;
}

/* ---------------------------------------------------------
   Crosstalk Checkbox Labels
   Makes the category labels next to checkbox filters slightly 
   larger for readability.
   --------------------------------------------------------- */

.crosstalk-input label {
  font-size: 14.5px !important;
}

/* ---------------------------------------------------------
   Global Font Size
   Overrides the default body font size for the whole dashboard,
   making all text slightly larger than the browser default (16px).
   --------------------------------------------------------- */

body {
  font-size: 18px !important;  
}

</style>


```{r setup, include=FALSE}
# ============================================================
# SETUP CHUNK
# 'include=FALSE' means this chunk runs silently — its code and
# output are NOT shown in the final dashboard.
#
# Here we:
#   1. Load all required R packages
#   2. Set a filename prefix used throughout the script
#   3. Apply a global option to suppress scientific notation
#   4. Apply a CSS rule to allow long column header text to wrap
# ============================================================

# --- Load packages ---
library(data.table)    # Fast data manipulation (used here mainly for compatibility)
library(DT)            # Creates interactive HTML data tables (DataTables.js)
library(patchwork)     # Combines multiple ggplot2 plots (loaded for compatibility)
library(plotly)        # Creates interactive, web-based plots
library(tidyverse)     # Core data science packages: dplyr, ggplot2, readr, tidyr, etc.
library(flexdashboard) # Turns R Markdown into an interactive web dashboard
library(crosstalk)     # Enables linked filtering across multiple HTML widgets
library(htmltools)     # Tools for building HTML elements from R

# --- Set filename prefix ---
prefix <- "VCFG_Data01"

# --- Suppress scientific notation ---
# This forces use of decimals to improve readability
options(scipen = 999)

# --- Allow DataTable column headers to wrap onto multiple lines ---
# By default, long column names in DT tables are truncated.
# This CSS injection allows header text to wrap so full names are visible.
tags$style(HTML("
  table.dataTable thead th {
    white-space: normal !important;
    word-break: break-word;
  }
"))
```


```{r input data}
# ============================================================
# DATA IMPORT
# Read in the single analysed data file from the project data folder.
# 
# This dataset has one row per well (i.e. per compound × dose),
# containing raw measurements, normalised values, robust Z-scores,
# and viability category assignments for each compound tested.
# ============================================================
data_analysed <- read_csv(paste0("../data/VCFG_Data01/", prefix, "_analysed_data_well.csv"))
```

```{r, prep data}
# ============================================================
# DATA PREPARATION
# Clean and restructure the imported data ready for display 
# and filtering in the dashboard.
# ============================================================

# --- Define ordered factor levels for Viability_Bin ---
# Factors with a defined order ensure that categories appear in a 
# meaningful sequence in filters and plots, rather than alphabetically.
vbin_order <- c("High", "Normal", "Moderate", "Low", "Very Low")

# --- Define the Z-score binning function ---
# Robust Z-scores can take a very wide range of values (e.g. -50 to +50).
# If used directly as slider filter values, this creates an unwieldy slider 
# with hundreds of tiny steps, making it hard to use.
#
# bin_z() solves this by:
#   1. Clamping extreme values to the range [-10, 10] with pmax/pmin
#      (any value below -10 becomes -10; any above 10 becomes 10)
#   2. Rounding to the nearest 0.5, with a 'dead zone' of 0 for values 
#      between -0.5 and +0.5 (treated as zero, i.e. near-baseline effect)
#   3. Using ceiling() for negative values and floor() for positive values
#      so that rounding always moves toward zero (conservative binning)
#
# The binned values are then used as the slider filter columns 
# (Bin_Feature_A_ZScore etc.) so the slider has manageable steps of 0.5.
# The original Z-scores are still shown in the table for accurate display.
bin_z <- function(x) {
  x <- pmax(pmin(x, 10), -10)        # Step 1: clamp to [-10, 10]
  ifelse(abs(x) < 0.5, 0,            # Step 2: values near zero become 0
         ifelse(x < 0, 
                ceiling(x / 0.5) * 0.5,   # Negative: round toward zero
                floor(x / 0.5) * 0.5))    # Positive: round toward zero
}

# --- Prepare the main dataframe ---
data_all <- data_analysed %>% 
  arrange(Compound_Name, desc(Concentration)) %>% 
  mutate(
    Concentration = paste0(Concentration, Units),
    Viability_Bin = factor(Viability_Bin, levels = vbin_order), # converting column to correct order as defined above
    
    # Apply the binning function to each Z-score column to create 
    # slider-friendly versions with steps of 0.5 (see bin_z() above).
    # These 'Bin_' columns are used ONLY for the slider filters — they 
    # are excluded from the download table.
    Bin_Feature_A_ZScore = bin_z(Robust_ZScore_Feature_A),
    Bin_Feature_B_ZScore = bin_z(Robust_ZScore_Feature_B),
    Bin_Feature_C_ZScore = bin_z(Robust_ZScore_Feature_C),
    
    # Create a unique identifier for each well by combining compound name 
    # and concentration. Because the same compound is tested at two doses,
    # compound name alone is not unique enough to identify a row.
    Unique_ID = paste(Compound_Name, Concentration, sep = "_"),
    
    # Eval_Key is the crosstalk linking key: widgets sharing the same group
    # use this column to match rows when a filter is applied.
    # Here it matches Unique_ID so filters act at the compound × dose level,
    # meaning selecting "Compound_X_10uM" does not also select "Compound_X_1uM".
    Eval_Key = paste(Compound_Name, Concentration, sep = "_")
  ) %>% 
  select(-Units)
```

```{r shared data, include=TRUE}
# ============================================================
# CROSSTALK SHARED DATA OBJECTS
#
# Crosstalk allows multiple HTML widgets (tables, sliders, checkboxes)
# to communicate with each other — when a filter is applied, all 
# widgets sharing the same group update simultaneously.
#
# SharedData$new() wraps a dataframe into a crosstalk-aware object.
# Key arguments:
#   group = "drug"   : All widgets in the same group are linked together.
#                      Filtering one widget filters all others in the group.
#   key = ~Eval_Key  : The column used to match rows across datasets.
#                      Rows with the same key value are highlighted/filtered 
#                      together. Here Eval_Key = Compound_Name + Concentration,
#                      so the 10uM and 1uM doses of the same compound are 
#                      treated as separate, independently-filterable entries.
#
# Two shared data objects are created from the same underlying data:
#
#   data_all_shared:    Includes the 'Bin_' columns, used by the slider 
#                       filters (which need to filter on binned Z-scores).
#
#   dt_data_all_shared: Excludes the 'Bin_' columns (via select(-starts_with("Bin_"))),
#                       used for the download table so users only see the 
#                       original Z-score values, not the internal binned versions.
#
# Both objects share the same group ("drug") and key (~Eval_Key), so 
# filtering via one also filters the other — the table always reflects 
# the current slider selections.
# ============================================================

# Full data including binned Z-score columns (for filter widgets)
data_all_shared <- data_all %>%
  SharedData$new(., group = "drug", key = ~Eval_Key)

# Data with binned columns removed (for the download table)
dt_data_all_shared <- data_all %>%
  select(-starts_with("Bin_")) %>%    # Drop all columns beginning with "Bin_"
  SharedData$new(., group = "drug", key = ~Eval_Key)
```

<!-- ============================================================
  DASHBOARD LAYOUT
  In flexdashboard, the layout is controlled by Markdown headings.
  
  '## Column' starts a new column panel.
  '### Panel Title' creates a named panel within a column.
  A column with no data-width attribute will fill the remaining space.
  ============================================================ -->

Row
-----------------------------------------------------------------------

### Primary Screen - Data Table

<!-- Static Markdown text displayed above the filters.
     Explains to the user how the filters work. -->

<br>
**Use the filters below to refine your hit list:**

* Slider bars: Move the handles on the sliders bars to select compounds with values within the desired range. The **purple** area of the slider bar indicates the value range you would like to keep.
* Cell viability tick boxes: Tick to select compounds in specific cell viability categories. The **ticked** boxes indicate the categories you would like to keep.

***Robust Z-score:***

```{r}
# ============================================================
# PANEL 1 FILTERS — Z-score Sliders (arranged side by side)
#
# Three sliders are created, one per feature's binned Z-score.
# They are placed inside a flexbox div so they appear in a 
# horizontal row rather than stacked vertically.
#
# div() from htmltools generates an HTML <div> element.
# The outer div uses CSS flexbox (display: flex) with a gap 
# between child divs. Each inner div holds one slider and has 
# a fixed width of 400px so sliders don't stretch unevenly.
#
# filter_slider() creates an interactive range slider. When the 
# user moves the handles, only rows with Bin values within the 
# selected range remain visible in all linked widgets.
#
# Arguments:
#   id          : A unique HTML ID for this filter element
#   label       : Text label displayed above the slider
#   sharedData  : The SharedData object this filter acts on.
#                 Note: uses data_all_shared (which contains the Bin_ columns)
#   ~Bin_...    : The binned Z-score column to filter on 
#   step        : The minimum increment the slider moves by (0.5, matching bin_z())
# ============================================================
div(
  style = "display: flex; gap: 40px; margin-bottom: 10px;",
  
  div(
    style = "flex: 0 0 400px;",   # Fixed width; won't shrink or grow
    filter_slider(
      id = "filter1",
      label = "Feature A:",
      sharedData = data_all_shared,
      ~Bin_Feature_A_ZScore,
      step = 0.5
    )
  ),
  
  div(
    style = "flex: 0 0 400px;",
    filter_slider(
      id = "filter2",
      label = "Feature B:",
      sharedData = data_all_shared,
      ~Bin_Feature_B_ZScore,
      step = 0.5
    )
  ),
  
  div(
    style = "flex: 0 0 400px;",
    filter_slider(
      id = "filter3",
      label = "Feature C:",
      sharedData = data_all_shared,
      ~Bin_Feature_C_ZScore,
      step = 0.5
    )
  )
)

# ============================================================
# PANEL 1 FILTER — Viability Bin Checkboxes
#
# filter_checkbox() creates a set of tick boxes, one per category.
# When boxes are ticked, only compounds in those categories are shown.
#
# Arguments:
#   allLevels = FALSE : Only show levels present in the current data
#                       (unlike TRUE, which would show all defined factor 
#                       levels even if currently filtered out)
#   inline = TRUE     : Place checkboxes side by side rather than stacked
#   columns = 1       : (with inline = TRUE) Arrange into 1 column
# ============================================================
filter_checkbox(
  id = "filter4",
  label = "Cell viability bins:",
  sharedData = data_all_shared,
  ~Viability_Bin,
  allLevels = FALSE,
  inline = TRUE,
  columns = 1
)
```

```{r}
# ============================================================
# RESET BUTTON
# A simple HTML button that reloads the entire page when clicked,
# which clears all filter selections back to their defaults.
# tags$button() is from htmltools — it generates a raw HTML <button>.
# ============================================================
tags$button(
  "Reset All Filters",
  class = "btn btn-default",        # Bootstrap button styling class
  onclick = "window.location.reload();",  # JavaScript: reload page on click
  style = "margin: 10px 0;"
)
```

```{r}
# ============================================================
# LIVE HIT COUNTER
#
# This block displays a running count of how many compounds pass 
# the current filters, split by concentration (10uM and 1uM).
# The counts update automatically whenever any filter changes.
#
# HOW IT WORKS:
# 
# 1. Two separate SharedData objects are created — one for high 
#    concentration (10uM) rows and one for low concentration (1uM).
#    Both share the same crosstalk group ("drug") and use Unique_ID 
#    as the key so they respond to the same filter events.
#    Only the Unique_ID column is included, since we only need the 
#    keys to count matches — no other data is required.
#
# 2. The initial counts and the full lists of keys are extracted from
#    these objects using $data() before any filtering occurs.
#    These are used to display the "X/total" format (e.g. "45/200").
#
# 3. browsable(tagList(...)) renders HTML elements directly in the 
#    dashboard. tagList() combines multiple HTML tags into one output.
#    browsable() tells R Markdown to render them as raw HTML.
#
# 4. The <div> block creates the visible counter display, with two 
#    <span> elements whose IDs ("high-count" and "low-count") are 
#    targeted by the JavaScript below to update the counts in real time.
#
# 5. The <script> block contains JavaScript that:
#    a. Converts the R key vectors to JavaScript Sets for fast lookup
#    b. Attaches a listener to the crosstalk FilterHandle for group "drug"
#    c. Each time a filter changes, the listener receives the list of 
#       keys that passed the filter (event.value)
#    d. It then counts how many of those keys belong to each concentration 
#       group and updates the text in the <span> elements
#    e. jsonlite::toJSON() converts the R vectors to JSON arrays 
#       that JavaScript can read (this runs at knit time, not in the browser)
# ============================================================

# --- Create minimal SharedData objects for counting ---
# Only Unique_ID is selected — we just need the keys for counting.
hits_shared_hi <- data_all %>%
  filter(Concentration == "10uM") %>%   # High concentration rows only
  select(Unique_ID) %>% 
  distinct() %>%                        # One row per unique compound × dose
  SharedData$new(., group = "drug", key = ~Unique_ID)

hits_shared_lo <- data_all %>%
  filter(Concentration == "1uM") %>%    # Low concentration rows only
  select(Unique_ID) %>% 
  distinct() %>% 
  SharedData$new(., group = "drug", key = ~Unique_ID)

# --- Extract initial totals and key lists ---
# $data() returns the underlying dataframe from a SharedData object
initial_high <- nrow(hits_shared_hi$data())   # Total number of 10uM compounds
initial_low  <- nrow(hits_shared_lo$data())   # Total number of 1uM compounds
high_keys    <- hits_shared_hi$data()$Unique_ID  # All 10uM compound keys
low_keys     <- hits_shared_lo$data()$Unique_ID  # All 1uM compound keys

# --- Build the counter UI ---
browsable(tagList(
  
  # The visible counter display panel
  tags$div(
    style = "margin-bottom: 20px; padding: 10px; background-color: #f0f0f0;",
    tags$div(
      tags$strong("High concentration compound matches: "),
      # This span's text is updated by the JS below when filters change
      tags$span(id = "high-count", paste0(initial_high, "/", initial_high))
    ),
    tags$div(
      tags$strong("Low concentration compound matches: "),
      tags$span(id = "low-count", paste0(initial_low, "/", initial_low))
    )
  ),
  
  # JavaScript that listens for filter changes and updates the counts
  tags$script(HTML(paste0("
    (function() {
      // Check crosstalk is available before running (it may not be in some environments)
      if (typeof crosstalk !== 'undefined') {
      
        // Convert R vectors to JavaScript Sets for fast membership testing.
        // jsonlite::toJSON() converts R vectors to JSON arrays at knit time,
        // which are then wrapped in 'new Set()' for O(1) lookup performance.
        var highKeys = new Set(", jsonlite::toJSON(high_keys), ");
        var lowKeys  = new Set(", jsonlite::toJSON(low_keys), ");
        var totalHigh = ", initial_high, ";
        var totalLow  = ", initial_low, ";
        
        // Attach to the 'drug' crosstalk group to listen for filter events
        var filterHandle = new crosstalk.FilterHandle('drug');
        
        // This function runs every time any filter in the 'drug' group changes.
        // event.value contains the list of keys that currently pass all filters.
        filterHandle.on('change', function(event) {
          var filteredKeys = event.value;
          var highCount, lowCount;
          
          // event.value === null means no filter is active — show all compounds
          if (filteredKeys === null) {
            highCount = totalHigh;
            lowCount  = totalLow;
            
          // event.value.length === 0 means a filter is active but nothing passes
          } else if (filteredKeys.length === 0) {
            highCount = 0;
            lowCount  = 0;
            
          // Otherwise, count how many of the passing keys belong to each group
          } else {
            var filteredSet = new Set(filteredKeys);
            highCount = 0;
            lowCount  = 0;
            
            // For each key that passed the filter, check which group it belongs to
            filteredSet.forEach(function(key) {
              if (highKeys.has(key)) highCount++;
              if (lowKeys.has(key))  lowCount++;
            });
          }
          
          // Update the text inside each <span> with the new X/total counts
          document.getElementById('high-count').textContent = highCount + '/' + totalHigh;
          document.getElementById('low-count').textContent  = lowCount  + '/' + totalLow;
        });
      }
    })();
  ")))
))
```


**Search for specific compounds**

```{r}
# ============================================================
# SEARCH DROPDOWNS — Compound Name and Concentration
#
# filter_select() creates a searchable dropdown menu.
# Unlike filter_slider() (which filters by value range), 
# filter_select() lets users search for and select specific 
# category values by name.
#
# Both dropdowns are placed in a horizontal row using a flexbox div,
# similar to the Z-score sliders above.
#
# Arguments:
#   multiple = TRUE  : Allow the user to select more than one value
#                      (e.g. search for several compounds at once)
#   allLevels = TRUE : Show all available options in the dropdown list,
#                      even if currently filtered out by other widgets
#
# These use data_all_shared (which includes the binned columns) but 
# they filter on Compound_Name and Concentration — the binned columns 
# are irrelevant to these dropdowns and are simply ignored.
# ============================================================
div(
  style = "display: flex; gap: 20px; margin-bottom: 20px;",

  filter_select(
    id = "filter5",
    label = "Filter by Compound Name",
    sharedData = data_all_shared,
    ~Compound_Name,
    allLevels = TRUE,
    multiple = TRUE    # Allow searching for multiple compounds at once
  ),

  filter_select(
    id = "filter6",
    label = "Filter by Concentration",
    sharedData = data_all_shared,
    ~Concentration,
    allLevels = TRUE,
    multiple = TRUE    # Allow selecting both concentrations simultaneously
  )
)
```


```{r}
# ============================================================
# MAIN DATA TABLE
#
# datatable() from the DT package creates an interactive HTML table
# with sorting, searching, pagination, and export buttons.
#
# Note: this table uses dt_data_all_shared (without the Bin_ columns)
# so users see the original Z-score values, not the internal binned
# versions used by the sliders. Because both shared data objects share
# the same crosstalk group and key, the table still responds to all 
# filter widgets correctly.
#
# Key arguments:
#   data            : The SharedData object to display (linked to filters)
#   extensions      : Extra DataTables plugins to enable
#     'Buttons'       : Adds Copy/Excel/column toggle buttons above the table
#     'FixedColumns'  : Freezes the leftmost columns during horizontal scrolling
#     'ColReorder'    : Lets the user drag-and-drop columns to reorder them
#   rownames = FALSE : Don't show row numbers in a separate first column
#   class            : CSS class applied to the table element
# ============================================================
datatable(
  dt_data_all_shared,   # Download table version (no Bin_ columns)
  extensions = c('Buttons', 'FixedColumns', 'ColReorder'),
  rownames = FALSE,
  escape = FALSE,       # Allow HTML formatting in cell content
  class = 'cell-border',
  options = list(
    
    # --- DOM layout string ---
    # Controls which DataTable UI elements appear and in what order.
    # B = Buttons, l = length menu (rows per page selector), s = search box, 
    # r = processing indicator, t = table, i = info text, p = pagination
    dom = 'Blsrtip',
    
    # --- drawCallback: JavaScript run each time the table redraws ---
    # This custom JS adds a thick purple border above the first row of 
    # each new compound group, making it visually clear when the compound 
    # changes in the table (since each compound has 2 rows: one per dose).
    drawCallback = JS(
      "function(settings) {",
      "  var api = this.api();",
      "",
      # Remove any borders added in the previous draw (prevents duplicates on redraw)
      "  api.rows({ page: 'current' }).nodes().each(function(row) {",
      "    $('td', row).css('border-top', '');",
      "  });",
      "",
      # Get the rows and data currently visible on this page
      "  var rows = api.rows({ page: 'current' }).nodes();",
      "  var data = api.rows({ page: 'current' }).data();",
      "",
      # Find the index of the Compound_Name column (needed for comparison below)
      "  var compoundIdx = api.column('Compound_Name:name').index();",
      "",
      # Loop through each row; if the compound name differs from the row above,
      # draw a thick purple top border — marking the start of a new compound
      "  for (var i = 0; i < data.length; i++) {",
      "    if (i === 0 || data[i][compoundIdx] !== data[i - 1][compoundIdx]) {",
      "      $('td', rows[i]).css('border-top', '2px solid #3c2372');",
      "    }",
      "  }",
      "}"
    ),
    
    # --- columnDefs: Controls the behaviour of specific columns ---
    columnDefs = list(
      # Centre-align all column content
      list(className = 'dt-center', targets = "_all"),
      
      # Give the Compound_Name column a named reference for the JS above.
      # which(colnames(...) == "Compound_Name") finds the zero-based column index.
      list(name = 'Compound_Name', targets = which(colnames(dt_data_all_shared) == "Compound_Name")),
      
      # Hide raw/normalised measurement columns by default — these are the 
      # per-replicate and averaged values from which Z-scores were calculated.
      # Users can reveal them via the 'Show/Hide Columns' button if needed.
      list(visible = FALSE, targets = c('Raw_A_Feature_A',                        
                                        'Raw_B_Feature_A',                       
                                        'Norm_A_Feature_A',                      
                                        'Norm_B_Feature_A',                       
                                        'Norm_Average_Feature_A',                 
                                        'Raw_A_Feature_B',              
                                        'Raw_B_Feature_B',              
                                        'Norm_A_Feature_B',            
                                        'Norm_B_Feature_B',             
                                        'Norm_Average_Feature_B',       
                                        'Raw_A_Feature_C',         
                                        'Raw_B_Feature_C',         
                                        'Norm_A_Feature_C',       
                                        'Norm_B_Feature_C',        
                                        'Norm_Average_Feature_C',  
                                        'Unique_ID',                                              
                                        'Eval_Key')),
      # Hide the linking columns completely — they are used for crosstalk 
      # filtering only and should never appear in the UI.
      # 'noVis' class also excludes them from the Show/Hide Columns menu.
      list(visible = FALSE, targets = c('Unique_ID', 'Eval_Key'), className = 'noVis')),
    
    pageLength = 20,   # Show 20 rows per page by default
    
    # --- Rows per page selector ---
    # Adds a dropdown above the table letting the user choose how many 
    # rows to display per page. The first list contains the numeric values;
    # the second list contains the labels shown in the dropdown.
    lengthMenu = list(c(20, 50, 100, 200), c('20', '50', '100', '200')),
    
    # --- Export/utility buttons ---
    buttons = list(
      # Copy visible table data to the clipboard as tab-separated text
      list(extend = 'copyHtml5',
           text = "Copy to Clipboard"),
      
      # Export visible table data to an Excel (.xlsx) file.
      # filename: the default name for the downloaded file
      # title = NULL: don't add a title row in the Excel file
      list(extend = 'excel',
           filename = paste0(prefix, "_Selected_Compounds"), 
           title = NULL),
      
      # A dropdown menu to show/hide individual columns.
      # columns = ':not(.noVis)' excludes the hidden crosstalk columns from the menu.
      list(extend = 'colvis',
           text = 'Show/Hide Columns',
           columns = ':not(.noVis)')),
    
    searching = TRUE,        # Show a text search box above the table
    scrollX = TRUE,          # Enable horizontal scrolling for wide tables
    scrollCollapse = TRUE,   # Don't add empty space below a short table
    autoWidth = FALSE,       # Don't auto-resize columns (use fixed widths)
    paging = TRUE,           # Break long tables into pages
    colReorder = TRUE,       # Allow users to reorder columns by dragging
    
    # Freeze the first 3 columns so they stay visible when scrolling right.
    # Useful when there are more columns than can fit on screen at once.
    fixedColumns = list(leftColumns = 3)),
  
  elementId = "data_download_table_1"   # A unique HTML ID for this table element
) %>% 
  
  # --- Conditional cell background colours ---

  # Alternating grey shading for the Concentration column.
  # Each dose gets a different shade so rows can be visually grouped by dose.
  # styleEqual() maps specific values to specific colours.
  # valueColumns = 'Concentration' means the colour is driven by the value 
  # in Concentration, even though the style is applied to both 
  # Compound_Name and Concentration columns together.
  formatStyle(
    c('Compound_Name', 'Concentration'),
    backgroundColor = styleEqual(
      c('10uM', '1uM'),
      c("#e9e9e9",    # Slightly darker grey for 10uM (higher dose)
        "#f7f7f7")    # Lighter grey for 1uM (lower dose)
    ),
    valueColumns = 'Concentration'
  ) %>%
  
  # Colour Z-score columns with a traffic-light scheme using styleInterval().
  # Unlike styleEqual() (which maps exact values), styleInterval() applies 
  # colours based on numeric thresholds (cut points):
  #   - Values BELOW the first threshold (-2) get the first colour (red)
  #   - Values BETWEEN the thresholds (-2 to +2) get the second colour (none)
  #   - Values ABOVE the last threshold (+2) get the third colour (green)
  # Note: styleInterval() always requires one more colour than threshold value.
  # 'transparent' means no background colour change for near-baseline Z-scores.
  formatStyle(
    columns = "Robust_ZScore_Feature_A",
    backgroundColor = styleInterval(
      c(-2, 2),                         # Cut points: below -2, between, above +2
      c("#fbeaea",                       # Z < -2: soft red (negative effect)
        "transparent",                   # -2 ≤ Z ≤ 2: no colour (baseline)
        "#e7f5e7")                       # Z > +2: soft green (positive effect)
    )
  ) %>%
  
  # Same colour scheme applied to Feature B Z-scores
  formatStyle(
    columns = "Robust_ZScore_Feature_B",
    backgroundColor = styleInterval(
      c(-2, 2),
      c("#fbeaea", "transparent", "#e7f5e7")
    )
  ) %>%
  
  # Same colour scheme applied to Feature C Z-scores
  formatStyle(
    columns = "Robust_ZScore_Feature_C",
    backgroundColor = styleInterval(
      c(-2, 2),
      c("#fbeaea", "transparent", "#e7f5e7")
    )
  ) %>%
  
  # Traffic-light colours for the Viability_Bin column using styleEqual().
  # Each named category gets a specific colour, ranging from red (very low 
  # viability — most toxic) to light blue (high viability — healthiest).
  formatStyle(
    'Viability_Bin',
    backgroundColor = styleEqual(
      c('Very Low', 'Low', 'Moderate', 'Normal', 'High'),
      c('#ffcdd2',    # Soft red (very low viability — most toxic)
        '#ffccbc',    # Soft orange
        '#fff9c4',    # Soft yellow
        '#f1f8e9',    # Soft green
        '#e1f5fe')    # Soft blue (high viability — healthy)
    )
  )
```
