---
# ============================================================
# YAML HEADER
# This block controls the overall appearance and layout of the 
# dashboard. It is not R code — it is read by the flexdashboard 
# package when knitting the document.
# ============================================================

title: "VCFG Dataset 4: Organoid Screen"   # Title shown in the navbar
output: 
  flexdashboard::flex_dashboard:
    theme:                      # Custom colour theme using Bootstrap Sass variables
      bg: "white"               # Background colour of the dashboard
      fg: "#3c2372"             # Foreground (text) colour — a dark purple
      primary: "#3c2372"        # Primary accent colour (buttons, active elements)
      secondary: "#3c2372"      # Secondary accent colour
      success: "#3c2372"        # Colour for "success" state elements
    orientation: columns        # Panels are arranged side-by-side in columns (vs rows)
    vertical_layout: fill       # Panels expand to fill the full height of the browser.
---

<!-- ============================================================
  CSS STYLING BLOCK
  Raw HTML/CSS injected directly into the dashboard. This customises
  the appearance of specific elements beyond what the YAML theme covers.
  Each section targets a different part of the UI.
  ============================================================ -->

<style type="text/css">

/* ---------------------------------------------------------
   DataTables Buttons (Copy, Excel, Show/Hide Columns)
   Makes the row of export buttons compact and left-aligned,
   rather than stretching to fill the available width.
   --------------------------------------------------------- */

.dataTables_wrapper .dt-buttons {
  display: inline-flex !important;  /* Lay buttons out in a horizontal row */
  align-items: center;
  gap: 6px;                         /* Small gap between each button */
  width: auto !important;           /* Don't stretch to full container width */
  flex: 0 0 auto !important;
}

.dt-buttons .dt-button {
  font-size: 13px !important;
  padding: 4px 8px !important;
  background: #f8f8f8 !important;   /* Light grey button background */
  border: 1px solid #bbb !important;
  border-radius: 4px !important;
}

.dt-buttons .dt-button:hover {
  background: #eaeaea !important;   /* Slightly darker on hover */
  border-color: #999 !important;
}

/* ---------------------------------------------------------
   Chart Panel Titles
   The text shown above each panel in the dashboard.
   --------------------------------------------------------- */

.chart-title {  
  font-size: 19px;
  font-weight: bold;
  color: #3c2372;   /* Match the dashboard's purple theme colour */
}

/* ---------------------------------------------------------
   ionRangeSlider (the interactive slider filter widgets)
   The slider is created by the crosstalk filter_slider() function.
   These rules customise its colours and hide default labels that
   would otherwise clutter the UI.
   --------------------------------------------------------- */

.irs-grid-text {
  color: #3c2372;
  display: none !important;   /* Hide the numeric grid labels beneath the slider */
}

.irs-bar {
  background-color: #3c2372; /* Purple fill for the selected range */
  border: 1px solid #3c2372;
}

.irs-slider {
  background: #3c2372 !important;   /* The draggable handle colour */
  border: 2px solid #3c2372;
  width: 10px !important;
  height: 10px !important;
  top: 25px !important;             /* Vertical position of the handle */
}

.irs-single,
.irs-from,
.irs-to {
  font-size: 14px !important;
  font-weight: bold;
  top: -4px !important;   /* Move the value label up slightly */
  color: #3c2372;
}

.irs-min,
.irs-max {
  display: none !important;   /* Hide the min/max endpoint labels */
}

/* ---------------------------------------------------------
   Navbar (the top navigation bar) — first block
   An earlier version of the navbar rules is present here;
   it is overridden by the more complete block below.
   --------------------------------------------------------- */

.navbar {
  min-height: 50px !important;
  height: 50px !important;
  padding-top: 8px !important;
  padding-bottom: 8px !important;
}

/* Align the navbar brand (the dashboard title) vertically within the bar */
.navbar-brand {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
  line-height: 20px !important;
  padding-top: 0 !important;
  padding-bottom: 0 !important;
  margin-top: 0 !important;
  margin-bottom: 0 !important;
  vertical-align: middle !important;
}

/* ---------------------------------------------------------
   Navbar (the top navigation bar) — full block
   Controls height and vertical alignment of all nav elements.
   This block supersedes the first .navbar block above.
   --------------------------------------------------------- */

.navbar {
  min-height: 50px !important;
  height: 50px !important;
  padding-top: 8px !important;
  padding-bottom: 8px !important;
}

/* Make nav items vertically centred inside the navbar */
.navbar-brand,
.navbar-nav,
.navbar-nav > li,
.navbar-header {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
}

.navbar-brand {
  line-height: 20px !important;
  padding: 0 !important;
  margin: 0 !important;
}

.navbar-nav {
  margin: 0 !important;
}

/* Style individual nav links — white text, no border, flex-aligned */
.navbar-nav > li > a {
  display: flex !important;
  align-items: center !important;
  height: 34px !important;
  line-height: 20px !important;
  padding: 7px 12px !important;
  background: none !important;
  border: none !important;
  color: white !important;
  font-weight: normal !important;
  text-decoration: none !important;
}

.navbar-nav > li > a:hover {
  color: rgba(255, 255, 255, 0.8) !important;  /* Slightly dimmed on hover */
}

/* Underline the currently active nav page */
.navbar-nav > li.active > a,
.navbar-nav > li > a.active,
.navbar-nav > li[class*="active"] > a {
  text-decoration: underline !important;
}

.navbar-nav > li.active > a:hover,
.navbar-nav > li > a.active:hover,
.navbar-nav > li[class*="active"] > a:hover {
  text-decoration: underline !important;
}

/* ---------------------------------------------------------
   Layout Spacing
   Reduces default padding/margins so the dashboard content
   fills more of the available screen space.
   --------------------------------------------------------- */

body {
  padding-top: 20px !important;   /* Gap below the navbar */
}

.container-fluid {
  padding-top: 10px !important;
  padding-left: 15px !important;
  padding-right: 15px !important;
  margin: 0 !important;
  width: 100% !important;   /* Use full browser width */
}

.row {
  margin-left: 0 !important;
  margin-right: 0 !important;
  width: 100% !important;
}

/* Tighten column gutters, with a small top padding */
.col-md-4,
.col-md-8 {
  padding-left: 5px !important;
  padding-right: 5px !important;
  padding-top: 10px !important;
}

/* Push content down to clear the fixed navbar height */
.navbar ~ .container-fluid,
.navbar + .container-fluid {
  margin-top: 55px !important;
}

/* Tighten bottom spacing on chart panels */
.section.level3,
.chart-wrapper,
.chart-stage,
.html-widget {
  margin-bottom: 5px !important;
  padding-bottom: 5px !important;
}

/* ---------------------------------------------------------
   Crosstalk Checkbox Labels
   Makes the category labels next to checkbox filters slightly 
   larger for readability.
   --------------------------------------------------------- */

.crosstalk-input label {
  font-size: 14.5px !important;
}

/* ---------------------------------------------------------
   Global Font Size
   Overrides the default body font size for the whole dashboard.
   A smaller size (14px) is used here compared to other dashboards
   because this dashboard has more content to display.
   --------------------------------------------------------- */

body {
  font-size: 14px !important;
}

</style>


```{r setup, include=FALSE}
# ============================================================
# SETUP CHUNK
# 'include=FALSE' means this chunk runs silently — its code and
# output are NOT shown in the final dashboard.
#
# Here we:
#   1. Load all required R packages
#   2. Set a filename prefix and cell line variable
#   3. Apply a global option to suppress scientific notation
#   4. Set global knitr chunk options
# ============================================================

# --- Load packages ---
library(data.table)    # Fast data manipulation (used here mainly for compatibility)
library(DT)            # Creates interactive HTML data tables (DataTables.js)
library(patchwork)     # Combines multiple ggplot2 plots (loaded for compatibility)
library(plotly)        # Creates interactive, web-based plots
library(tidyverse)     # Core data science packages: dplyr, ggplot2, readr, tidyr, etc.
library(flexdashboard) # Turns R Markdown into an interactive web dashboard
library(crosstalk)     # Enables linked filtering across multiple HTML widgets
library(htmltools)     # Tools for building HTML elements from R
library(htmlwidgets)   # Framework for creating R bindings to JavaScript libraries;
                       # used here for onRender() to run custom JS on a DT widget
library(purrr)         # Functional programming tools (map, walk, etc.); loaded
                       # for compatibility though lapply() is used directly below

# --- Set filename prefix ---
prefix <- "VCFG_Data04"

# --- Suppress scientific notation ---
# This forces use of decimals to improve readability
options(scipen = 999)

# --- Set global knitr chunk options ---
# These apply to all subsequent code chunks unless overridden locally.
# echo = FALSE: don't show any R code in the knitted output (just results)
# collapse = FALSE: keep code output separate from the code itself (default behaviour)
knitr::opts_chunk$set(
  echo = FALSE,
  collapse = FALSE
)
```

```{r cell line original data, include=TRUE}
# ============================================================
# DATA IMPORT AND PREPARATION — CELL LINE DATA
#
# This dashboard works with data from a single cell line (c).
# Three separate data files are read in and processed:
#
#   1. pred_all_data_mean:  Per-well imaging model predictions 
#                           (average probability a cell is alive)
#   2. pred_all_data_aucn:  Per-compound imaging model summary 
#                           (AUC of survival probability across doses)
#   3. analysed_data_comp:  Per-compound CTG viability summary stats
#                           (CTG AUC, Robust Z-scores)
#   4. analysed_data_well:  Per-well CTG viability measurements
#
# The imaging data (pred_all_data_mean, pred_all_data_aucn) and 
# CTG data (analysed_data_comp, analysed_data_well) are joined 
# together so each compound/well has both CTG and ML predictions.
# ============================================================

# --- Set the cell line ---
c <- "Cell_LineX"

# ============================================================
# FILE 1: Per-well imaging prediction data
# Contains one row per well with the model's predicted probability
# that cells in that well are alive (prob_alive: 0 = dead, 1 = alive).
# ============================================================
pred_all_data_mean <- read_csv(paste0("../data/VCFG_Data04/", prefix, "_pred_all_data_mean.csv")) %>% 
  select(VCFG_Plate_ID = Plate_Folder,  
         WellID, 
         prob_alive) %>% 
  mutate_at(3, round, 2)                

# ============================================================
# FILE 2: Per-compound imaging prediction summary
# Contains one row per compound with AUC metrics from the ML model.
# ML_AUC is the raw area under the survival curve; ML_Relative_AUC 
# is that value normalised to the DMSO control.
# ============================================================
pred_all_data_aucn <- read_csv(paste0("../data/VCFG_Data04/", prefix, "_pred_all_data_aucn.csv")) %>% 
  select(Compound_Name, 
         ML_AUC = AUC,                   
         ML_Relative_AUC = Relative_AUC) %>% 
  mutate_at(2:3, round, 2)             

# ============================================================
# FILE 3: Per-compound CTG viability summary
# Contains one row per compound with AUC and Z-score metrics
# calculated from CTG (CellTiter-Glo) luminescence viability data.
# ============================================================
analysed_data_comp <- read_csv(paste0("../data/VCFG_Data04/", prefix, "_analysed_data_comp.csv")) %>%
  filter(Cell_Line == c) %>%
  left_join(., pred_all_data_aucn, by = "Compound_Name") %>% 
  
  mutate(
    Direction = case_when(
      Robust_ZScore_AUC < 0  ~ "Negative",
      Robust_ZScore_AUC >= 0 ~ "Positive",
      TRUE ~ NA),
    ZScore_abs = abs(Robust_ZScore_AUC),
    Robust_ZScore_Bin = case_when(
      ZScore_abs >= 0   & ZScore_abs < 2   ~ 0,
      ZScore_abs >= 2   & ZScore_abs < 2.5 ~ 2,
      ZScore_abs >= 2.5 & ZScore_abs < 3   ~ 2.5,
      ZScore_abs >= 3   & ZScore_abs < 3.5 ~ 3,
      ZScore_abs >= 3.5 & ZScore_abs < 4   ~ 3.5,
      ZScore_abs >= 4   & ZScore_abs < 4.5 ~ 4,
      ZScore_abs >= 4.5 & ZScore_abs < 5   ~ 4.5,
      ZScore_abs >= 5   & ZScore_abs < 5.5 ~ 5,
      ZScore_abs >= 5.5 & ZScore_abs < 6   ~ 5.5,
      ZScore_abs >= 6   & ZScore_abs < 6.5 ~ 6,
      ZScore_abs >= 6.5 & ZScore_abs < 7   ~ 6.5,
      ZScore_abs >= 7   & ZScore_abs < 7.5 ~ 7,
      ZScore_abs >= 7.5 & ZScore_abs < 8   ~ 7.5,
      ZScore_abs >= 8                       ~ 8,
      TRUE ~ NA)
  )

# ============================================================
# FILE 4: Per-well CTG viability data
# Contains one row per well (compound × dose) with raw and
# normalised CTG luminescence values and their Z-scores.
# ============================================================
analysed_data_well <- read_csv(paste0("../data/VCFG_Data04/", prefix, "_analysed_data_well.csv")) %>% 
  filter(Cell_Line == c) %>% 
  left_join(., pred_all_data_mean, by = c("VCFG_Plate_ID", "WellID")) %>%
  mutate(
    Conc = paste0(Concentration, Units),
    ZScore_Sig = case_when(
      Robust_ZScore <= -2 | Robust_ZScore >= 2 ~ "Significant",
      TRUE ~ "Not Significant")
  ) %>% 
  mutate(
    ML_Viability_Bin = case_when(
      prob_alive >= 0   & prob_alive < 0.2 ~ "Very Low",
      prob_alive >= 0.2 & prob_alive < 0.4 ~ "Low",
      prob_alive >= 0.4 & prob_alive < 0.7 ~ "Moderate",
      prob_alive >= 0.7 & prob_alive < 1.1 ~ "Normal",
      TRUE ~ NA)
  )

# ============================================================
# WIDE FORMAT DATA — For per-dose-point checkbox filters
# ============================================================
analysed_data_well_wide <- analysed_data_well %>% 
  pivot_wider(
    id_cols = c(Compound_Name),
    names_from = Conc,                            # Creates column names like ZScore_Sig_10uM
    values_from = c(ZScore_Sig, ML_Viability_Bin) # Two sets of new columns, one per dose
  ) %>% 
  mutate(
    across(everything(), ~ifelse(is.na(.), "Missing Data", .)),
    across(starts_with("ZScore_Sig"), 
           ~factor(., levels = c("Missing Data", "Not Significant", "Significant"))),
    across(starts_with("ML_Viability"),
           ~factor(., levels = c("Missing Data", "Very Low", "Low", "Moderate", "Normal")))
  )
```

```{r cell line display data, include=TRUE}
# ============================================================
# DISPLAY DATA — Trimmed tables for the dashboard tables
# disp_data_comp: Compound-level summary (one row per compound)
# disp_data_well: Dose-level measurements (one row per well)
# ============================================================

disp_data_comp <- analysed_data_comp %>% 
  select(Compound_Name,
         Relative_AUC, Robust_ZScore_AUC,   # CTG-based metrics
         ML_AUC, ML_Relative_AUC) %>%        # ML imaging-based metrics
  arrange(Compound_Name)

disp_data_well <- analysed_data_well %>%
  select(Compound_Name, Conc,
         CTG, CTG_Norm, Robust_ZScore,       # Raw and normalised CTG values
         prob_alive, ML_Viability_Bin) %>%   # ML imaging prediction columns
  arrange(Compound_Name)
```

```{r cell line shared data, include=TRUE}
# ============================================================
# CROSSTALK SHARED DATA OBJECTS
#
# Crosstalk allows multiple HTML widgets (tables, sliders, checkboxes,
# plots) to communicate with each other — when a filter is applied,
# all widgets sharing the same group update simultaneously.
#
# SharedData$new() wraps a dataframe into a crosstalk-aware object.
# Key arguments:
#   group = "drug"        : All widgets in this group are linked together.
#                           Filtering one widget filters all others in the group.
#   key = ~Compound_Name  : The column used to match rows across datasets.
#                           ALL shared data objects here use Compound_Name as 
#                           the key so that selecting/filtering a compound updates
#                           every linked widget simultaneously.
#
# FIVE shared data objects are created, serving different roles:
#
#   data_comp_shared:    Full compound-level data (for filter widgets + plot).
#                        Includes the binned Z-score and Direction columns 
#                        needed by the sidebar filters.
#
#   data_well_shared:    Full well-level data (for filter widgets).
#                        Contains the raw per-dose measurements; filters applied 
#                        here propagate to the dose-level display table.
#
#   data_well_wide_shared: Wide-format well data (for per-dose checkbox filters).
#                        This is the pivoted data with one row per compound and 
#                        separate columns per dose (e.g. ZScore_Sig_10uM).
#                        Checkbox filters for individual dose points are linked 
#                        to this object because crosstalk checkboxes require a 
#                        single column per filter, not long-format data.
#
#   disp_comp_shared:    Trimmed compound display data (for the compound table).
#                        Only contains columns shown to the user; shares the same 
#                        group so the table responds to all sidebar filters.
#
#   disp_well_shared:    Trimmed well display data (for the dose-level table).
#                        Only contains columns shown to the user; responds to the 
#                        same filters as the compound table.
# ============================================================

# Full data for filter widgets and plot
data_comp_shared     <- SharedData$new(analysed_data_comp,      group = "drug", key = ~Compound_Name)
data_well_shared     <- SharedData$new(analysed_data_well,      group = "drug", key = ~Compound_Name)
data_well_wide_shared <- SharedData$new(analysed_data_well_wide, group = "drug", key = ~Compound_Name)

# Trimmed data for display/download tables
disp_comp_shared <- SharedData$new(disp_data_comp, group = "drug", key = ~Compound_Name)
disp_well_shared <- SharedData$new(disp_data_well, group = "drug", key = ~Compound_Name)
```

<!-- ============================================================
  DASHBOARD LAYOUT
  In flexdashboard, the layout is controlled by Markdown headings.
  
  '## Column {data-width=350}' starts a new column with a fixed 
  pixel width of 350. The remaining columns share the leftover space.
  '### Panel Title' creates a named panel within a column.
  ============================================================ -->

Column {data-width=350}
-----------------------------------------------------------------------

### Hit selection filters

<!-- Static Markdown text explaining the filter sections. -->

***1. Filter compounds based on the CTG Robust Z-score across the entire dose-range (calculated based on the relative AUC)*** <br>
Z-scores ±2 or more are generally considered to be significant, but this is dependent on the hit rate and strength of the equivalent fold change values.

```{r}
# ============================================================
# FILTER 1 — Z-score Direction Checkbox
#
# This checkbox filter lets users select whether to view compounds 
# with negative Z-scores (decreased viability), positive Z-scores 
# (increased viability), or both.
#
# The Direction column in data_comp_shared was pre-calculated in the 
# data prep step as "Negative" or "Positive" based on Robust_ZScore_AUC.
#
# Note: the JavaScript block below modifies this filter on page load 
# to pre-select "Negative" only and enforce single-selection behaviour,
# since this screen is primarily searching for compounds that reduce viability.
#
# allLevels = TRUE: Show all factor levels, even if currently filtered out
# inline = TRUE:    Place checkboxes side by side rather than stacked
# columns = 1:      With inline = TRUE, arrange into a single column
# ============================================================
filter_checkbox(
  id = "filter1",
  label = "Negative values indicate a decrease, positive values indicate an increase.",
  sharedData = data_comp_shared,
  ~Direction,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)
```

```{r}
# ============================================================
# FILTER 2 — Z-score Magnitude Slider
#
# This slider filters compounds by the absolute magnitude of their 
# Z-score. Combined with the Direction filter above, it allows the 
# user to select compounds with, for example, a Z-score more negative 
# than -3 (Direction = Negative; slider min = 3).
#
# The slider uses Robust_ZScore_Bin (the binned absolute Z-score from 
# data prep) rather than the raw Z-score, so the slider moves in 
# clean 0.5-unit steps rather than through many decimal place values.
#
# round = FALSE: Don't round the slider value to the nearest integer
# ticks = FALSE: Don't show tick marks below the slider track
# ============================================================
filter_slider(
  id = "filter2",
  label = "Z-score cut-off:",
  sharedData = data_comp_shared,
  ~Robust_ZScore_Bin,
  step = 0.5,
  round = FALSE,
  ticks = FALSE
)
```

<br>

***2. Filter compounds based on acceptable CTG Z-scores at individual dose points*** <br>
Un-tick the *Not Significant* option to filter for compounds with a significant CTG Z-score (±2) at each dose-point. 

```{r}
# ============================================================
# FILTERS 3–5 — Per-dose-point Z-score Significance Checkboxes
#
# These three checkboxes filter at the individual dose level, letting 
# users require that a compound has a significant Z-score at a specific 
# concentration (10µM, 1µM, or 0.1µM).
#
# They are linked to data_well_wide_shared (the pivoted wide-format data)
# because each dose has its own column (ZScore_Sig_10uM, ZScore_Sig_1uM, 
# ZScore_Sig_0.1uM). Crosstalk checkboxes require a single column per 
# filter — the wide format provides this.
#
# Each column has three possible values: "Missing Data", "Not Significant",
# and "Significant". The allLevels = TRUE argument ensures all three 
# options always appear even if some are currently filtered out.
#
# The JavaScript block below pre-ticks all options by default (i.e. no 
# compounds are excluded by these filters until the user actively un-ticks).
# ============================================================
filter_checkbox(
  id = "filter3",
  label = "10μM",
  sharedData = data_well_wide_shared,
  ~ZScore_Sig_10uM,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)

filter_checkbox(
  id = "filter4",
  label = "1μM",
  sharedData = data_well_wide_shared,
  ~ZScore_Sig_1uM,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)

filter_checkbox(
  id = "filter5",
  label = "0.1μM",
  sharedData = data_well_wide_shared,
  ~ZScore_Sig_0.1uM,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)
```

<br>

***3. Filter compounds based on ML predicted viability at individual dose points*** <br>
Un-tick the *Moderate* and/or *Normal* options to filter for compounds with very-low/low predicted cell viability at each dose-point.

```{r}
# ============================================================
# FILTERS 6–8 — Per-dose-point ML Viability Checkboxes
#
# These three checkboxes work the same way as filters 3–5, but 
# filter on the ML model's predicted viability category at each dose 
# (ML_Viability_Bin_10uM, ML_Viability_Bin_1uM, ML_Viability_Bin_0.1uM)
# rather than on CTG Z-score significance.
#
# The four possible viability categories are:
#   "Missing Data", "Very Low", "Low", "Moderate", "Normal"
# Un-ticking "Moderate" and "Normal" restricts the view to compounds 
# where the ML model predicts notable cell death at that dose.
# ============================================================
filter_checkbox(
  id = "filter6",
  label = "10μM",
  sharedData = data_well_wide_shared,
  ~ML_Viability_Bin_10uM,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)

filter_checkbox(
  id = "filter7",
  label = "1μM",
  sharedData = data_well_wide_shared,
  ~ML_Viability_Bin_1uM,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)

filter_checkbox(
  id = "filter8",
  label = "0.1μM",
  sharedData = data_well_wide_shared,
  ~ML_Viability_Bin_0.1uM,
  allLevels = TRUE,
  inline = TRUE,
  columns = 1
)
```

```{js}
// ============================================================
// JAVASCRIPT: Default filter state and custom filtering logic
//
// This block runs once after the page has fully loaded. It handles
// three separate tasks:
//
//   1. Pre-select "Negative" in filter1 and enforce single-selection
//   2. Pre-tick all options in filters 3–8 (so no rows are excluded by default)
//   3. Set up cross-table synchronisation between the compound and well tables
//
// $(document).ready() runs the code inside once the page DOM is loaded.
// setTimeout(..., 300) adds a 300ms delay to ensure all crosstalk widgets 
// have fully initialised before the script tries to interact with them.
// ============================================================
$(document).ready(function() {
  setTimeout(function() {

    // -------- TASK 1: Set up filter1 (Direction checkbox) --------
    // This screen is looking for compounds that reduce viability, so 
    // the page should load with "Negative" pre-selected.
    // We also enforce single-selection so only one direction can be 
    // active at a time.

    // First, uncheck ALL checkboxes in filter1
    $("#filter1 input[type='checkbox']").prop("checked", false);

    // Then check only the "Negative" option and trigger a change event
    // so crosstalk registers the new selection
    $("#filter1 input[type='checkbox'][value='Negative']").prop("checked", true).trigger("change");

    // Add an event handler so that checking one option unchecks all others
    $("#filter1 input[type='checkbox']").on("change", function() {
      var $this = $(this);

      if ($this.is(":checked")) {
        // Uncheck all other boxes and trigger their change events
        $("#filter1 input[type='checkbox']").not($this).prop("checked", false).trigger("change");
      } else {
        // Prevent the user from unchecking the last box (at least one must be selected)
        var anyChecked = $("#filter1 input[type='checkbox']:checked").length > 0;
        if (!anyChecked) {
          $this.prop("checked", true);
        }
      }
    });

    // -------- TASK 2: Pre-tick all options in filters 3–8 --------
    // By default, crosstalk checkboxes load with nothing ticked, which 
    // hides all data. We pre-tick everything so all compounds are visible 
    // on page load, and users only need to un-tick to restrict the view.
    // .trigger("change") fires the crosstalk change event so the tables update.
    $("#filter3 input[type='checkbox']").prop("checked", true).trigger("change");
    $("#filter4 input[type='checkbox']").prop("checked", true).trigger("change");
    $("#filter5 input[type='checkbox']").prop("checked", true).trigger("change");
    $("#filter6 input[type='checkbox']").prop("checked", true).trigger("change");
    $("#filter7 input[type='checkbox']").prop("checked", true).trigger("change");
    $("#filter8 input[type='checkbox']").prop("checked", true).trigger("change");

    // -------- TASK 3: Cross-table synchronisation --------
    // The compound-level table and the dose-level well table are both 
    // linked to the same crosstalk group ("drug"), so filtering one 
    // should filter the other. However, crosstalk's built-in linking 
    // requires each dataset to share an identical key column — and here, 
    // the compound table has one row per compound while the well table 
    // has multiple rows per compound (one per dose).
    //
    // The custom functions below work around this by:
    //   a. Reading the currently visible compound names from the compound table
    //   b. Manually filtering the well table to show only rows matching 
    //      those compound names
    //
    // This ensures that filtering by compound-level metrics (Z-score, direction)
    // also correctly updates the per-dose well table.

    // --- Helper: find which compound names are currently passing filters ---
    function getVisibleCompounds() {
      var visibleCompounds = [];
      // Find the compound table widget by looking for 'compound' in its ID
      // or 'Per Compound' in the heading of its containing section
      var compoundTable = window.HTMLWidgets.find(".datatables").filter(function(widget) {
        return widget.el.id.includes("compound") || $(widget.el).closest('.section').find('h3').text().includes('Per Compound');
      })[0];
      
      if (compoundTable && compoundTable.x && compoundTable.x.data) {
        var filteredData = compoundTable.x.crosstalkFilter.filteredKeys;
        if (filteredData && filteredData.length > 0) {
          visibleCompounds = filteredData;   // Use the crosstalk-filtered key list
        } else {
          // No active filter — collect all compound names from the raw table data
          compoundTable.x.data.forEach(function(row, index) {
            var compoundName = row[0];  // Column 0 = Compound_Name
            if (compoundName) visibleCompounds.push(compoundName);
          });
        }
      }
      return visibleCompounds;
    }
    
    // --- Helper: update the well table to show only matching compounds ---
    function updateWellTable() {
      var visibleCompounds = getVisibleCompounds();
      
      // Find the well table widget by looking for 'well' in its ID
      // or 'Per Dose' in the heading of its containing section
      var wellTable = window.HTMLWidgets.find(".datatables").filter(function(widget) {
        return widget.el.id.includes("well") || $(widget.el).closest('.section').find('h3').text().includes('Per Dose');
      })[0];
      
      if (wellTable && wellTable.x && wellTable.x.data) {
        var visibleRows = [];
        
        // Check each row of the well table; keep those whose compound is in the visible list
        wellTable.x.data.forEach(function(row, index) {
          var compoundName = row[0];  // Column 0 = Compound_Name
          if (visibleCompounds.includes(compoundName)) {
            visibleRows.push(index);
          }
        });
        
        // Push the updated row visibility to the well table's crosstalk filter
        if (wellTable.x.crosstalkFilter) {
          wellTable.x.crosstalkFilter.set(visibleRows);
        }
      }
    }
    
    // Listen for any crosstalk filter change event and update the well table.
    // The 100ms delay ensures the compound table has finished updating before 
    // we try to read its visible rows.
    $(document).on('crosstalk:filter', function(event) {
      setTimeout(updateWellTable, 100);
    });
    
    // Run once on page load so the well table is correct from the start
    setTimeout(updateWellTable, 1000);

  }, 300);
});
```

Column 
-----------------------------------------------------------------------

### Compound-Level Table

```{r}
# ============================================================
# COMPOUND-LEVEL DATA TABLE
#
# datatable() creates an interactive HTML table. This table is 
# additionally extended with a custom onRender() callback (from 
# the htmlwidgets package) that connects table row clicks to the 
# image panel in the third column — clicking a row in this table 
# updates the compound images displayed below the scatter plot.
#
# Extensions used:
#   'Buttons'  : Adds Copy/Excel buttons above the table
#   'Scroller' : Replaces pagination with virtual scrolling — the table 
#                scrolls smoothly through all rows without page breaks,
#                only rendering the visible rows for performance
#   'FixedColumns': Freezes the leftmost columns during horizontal scrolling
# ============================================================
comp_dt <- datatable(
  disp_comp_shared, 
  extensions = c('Buttons', 'Scroller', "FixedColumns"), 
  rownames = FALSE,
  options = list(
    dom = 'Brfti',   # B=Buttons, r=processing, f=filter/search, t=table, i=info text
    columnDefs = list(list(className = 'dt-center', targets = "_all")),
    pageLength = 50,
    buttons = list(
      list(extend = 'copyHtml5', text = "Copy to Clipboard"),
      list(extend = 'excel',
           filename = paste0(prefix, "_Hits_PerCompound_Data_", c), 
           title = NULL)),
    searching = TRUE,
    scrollX = TRUE,
    paging = TRUE,
    fixedColumns = list(leftColumns = 2)
  )
) %>%
  
  # ============================================================
  # onRender() — Custom JavaScript run when the table is first drawn
  #
  # This callback is from the htmlwidgets package. It runs client-side 
  # JavaScript after the table widget is rendered in the browser, giving 
  # access to the underlying DataTables API (tbl) and the DOM element (el).
  #
  # The code here does two things:
  #   1. On page load, pushes the first visible compound name to 
  #      window.updateCompoundImages() so the image panel shows 
  #      relevant images immediately.
  #   2. Adds a click handler to table rows so that clicking any row 
  #      calls window.updateCompoundImages() with that compound's name,
  #      updating the images in the panel below the plot.
  #   3. On table redraw (e.g. after sorting or searching), updates the 
  #      images to match the new first visible row.
  #
  # window.updateCompoundImages is defined in the image panel chunk below — 
  # it is attached to the global window object so it can be called from 
  # anywhere in the page's JavaScript.
  # ============================================================
  htmlwidgets::onRender("
    function(el, x) {
      var $el = $(el); 
      var tbl = $el.find('table').DataTable();

      var compoundCol = 0; // 0-based column index: column 0 = Compound_Name

      // Helper: push a compound name to the image updater
      function pushCompound(comp) {
        if (window.updateCompoundImages) window.updateCompoundImages(comp);
      }

      // On initial render, show images for the first visible row
      var first = tbl.row({search:'applied'}).data();
      if (first && first[compoundCol]) pushCompound(first[compoundCol]);

      // On row click: update images to match the clicked compound
      $el.on('click', 'tbody tr', function() {
        var data = tbl.row(this).data();
        if (!data) return;
        var comp = data[compoundCol];
        if (!comp) return;
        pushCompound(comp);
      });

      // After any table redraw (sort, search, filter): update to new first row
      tbl.on('draw', function() {
        var d = tbl.row({search:'applied'}).data();
        if (d && d[compoundCol]) pushCompound(d[compoundCol]);
      });
    }
  ")

# Render the table in the dashboard panel
comp_dt
```

### Dose-Point-Level Table

```{r}
# ============================================================
# DOSE-LEVEL DATA TABLE
#
# Displays one row per compound × dose, showing CTG and ML metrics 
# for each individual dose point. This table responds to the same 
# crosstalk filters as the compound table via the shared "drug" group.
#
# Uses the same extensions as the compound table but with 3 fixed 
# columns (compound name + dose are needed for context when scrolling).
# No onRender() callback here as this table doesn't control the images.
# ============================================================
datatable(
  disp_well_shared, 
  extensions = c('Buttons', 'Scroller', "FixedColumns"), 
  rownames = FALSE,
  options = list(
    dom = 'Brft',   # No 'i' info text (row count) to save vertical space
    columnDefs = list(list(className = 'dt-center', targets = "_all")),
    pageLength = 50,
    buttons = list(
      list(extend = 'copyHtml5', text = "Copy to Clipboard"),
      list(extend = 'excel',
           filename = paste0(prefix, "_Hits_PerWell_Data_", c), 
           title = NULL)),
    searching = TRUE,
    scrollX = TRUE,
    paging = TRUE,
    fixedColumns = list(leftColumns = 3)   # Fix Compound_Name, Conc, and one more column
  )
)
```

Column
-----------------------------------------------------------------------

### Compound-Level Plot

```{r}
# ============================================================
# SCATTER PLOT — ML Relative AUC vs CTG Robust Z-score
#
# This plot shows each compound as a single point, with:
#   x-axis: ML Relative AUC (machine learning predicted viability 
#            as a fold-change to DMSO control; 1 = no effect)
#   y-axis: CTG AUC Robust Z-score (statistical measure of how 
#            different the CTG viability is from the plate median)
#
# Points are coloured by compound name. Because the number of 
# compounds is typically large, the legend is present but the 
# main purpose of the colour is to help track individual compounds 
# when hovering or after filtering, not to distinguish all at once.
#
# The plot is linked to data_comp_shared so it responds to the 
# sidebar filters — compounds filtered out are hidden from the plot.
# ============================================================

# --- Pull the first compound name (for the image panel default) ---
first_compound <- analysed_data_comp %>%
  dplyr::pull(Compound_Name) %>%
  unique() %>%
  .[1]   # .[1] extracts the first element of the vector

# --- Calculate dynamic axis limits ---
# The y-axis (Z-score) limits are calculated from the data rather than 
# hard-coded, ensuring the plot scales correctly regardless of the 
# range of values in any particular dataset.
#
# The minimum y-axis limit is at least -5 (even if no Z-score is below -5),
# and the maximum is at least +5. This provides a reasonable baseline scale
# while expanding further if the data requires it.
# floor() and ceiling() round down/up to the nearest integer.
max_lim_AUC_z <- analysed_data_comp %>% 
  summarise(value = max(Robust_ZScore_AUC, na.rm = TRUE)) %>% 
  mutate(value = if_else(value < 5, 5, value + 1),  
         value = ceiling(value)) %>%                 
  pull(value)

min_lim_AUC_z <- analysed_data_comp %>% 
  summarise(value = min(Robust_ZScore_AUC, na.rm = TRUE)) %>% 
  mutate(value = if_else(value > -5, -5, value + 1),  
         value = floor(value)) %>%                      
  pull(value)

# Fixed x-axis limits for ML Relative AUC (0 = no cells, 1 = DMSO-equivalent)
max_lim_rAUC <- 1.2
min_lim_rAUC <- 0

# --- Build the scatter plot ---
comp_dot_plotly <- data_comp_shared %>%
  plot_ly(., 
          x = ~ML_Relative_AUC,       # x-axis: ML predicted viability
          y = ~Robust_ZScore_AUC,     # y-axis: CTG Z-score
          color = ~Compound_Name,     # One colour per compound
          legendgroup = ~Compound_Name,
          text = ~Compound_Name,
          type = 'scatter',
          mode = 'markers',
          marker = list(opacity = 0.8),  # Slight transparency helps with overlapping points
          hovertemplate = ~paste0(
            "<b>", Compound_Name, "</b>",
            '<br>CTG AUC Robust Z-score: ', Robust_ZScore_AUC,
            '<br>CTG Relative AUC: ', Relative_AUC, 
            '<br>ML Relative AUC: ', ML_Relative_AUC,
            "<extra></extra>"   # Suppresses the default plotly trace name box
          )) %>%
  
  # --- Axis and plot layout ---
  layout(
    xaxis = list(
      title = "ML predicted viability AUC (Fold Change to DMSO)", 
      titlefont = list(size = 13.5),
      range = c(min_lim_rAUC, max_lim_rAUC),
      tickvals = seq(min_lim_rAUC, max_lim_rAUC, by = 0.5),
      showline = TRUE,
      zeroline = FALSE
    ),
    yaxis = list(
      title = "AUC Robust Z-score",
      titlefont = list(size = 13.5),
      range = c(min_lim_AUC_z, max_lim_AUC_z),
      tickvals = seq(from = ceiling(min_lim_AUC_z / 5) * 5,
                     to   = floor(max_lim_AUC_z / 5) * 5,
                     by   = 5),
      showline = TRUE
    ),
    legend = list(title = list(text = "Colour: Compound\n")),
    margin = list(t = 30, r = 40, b = 65, l = 65),
    plot_bgcolor = "white"   
  ) %>% 
  
  # --- Reference lines ---
  # Added as layout 'shapes' so they appear as fixed lines behind the data,
  # without creating legend entries.
  layout(shapes = list(
    list(type = "line",
         x0 = 1, x1 = 1,
         y0 = min_lim_AUC_z, y1 = max_lim_AUC_z,
         line = list(dash = "dot", color = "black", width = 0.6)),
    list(type = "line",
         x0 = min_lim_rAUC, x1 = max_lim_rAUC,
         y0 = -2, y1 = -2,
         line = list(dash = "dot", color = "black", width = 0.6)),
    list(type = "line",
         x0 = min_lim_rAUC, x1 = max_lim_rAUC,
         y0 = 2, y1 = 2,
         line = list(dash = "dot", color = "black", width = 0.6))
  )) 

comp_dot_plotly
```

### Endpoint Images

<!-- Static Markdown text explaining the image panel. -->

Use the plot above to select a compound of interest. Double-click on the data point and the corresponding brightfield and DAPI images for each concentration will be shown below. The 0.2% DMSO control is from plate 1 well B16.

```{r}
# ============================================================
# DYNAMIC IMAGE PANEL
#
# This panel displays microscopy images for a selected compound.
# It has two sections:
#
#   LEFT:  Two static DMSO control images (brightfield + DAPI), 
#          which never change — they are always the same reference images.
#
#   RIGHT: Six dynamic compound images (brightfield + DAPI at three 
#          doses: 0.1µM, 1µM, 10µM), which update when the user clicks 
#          a compound in the table above or double-clicks the plot.
#
# HOW THE IMAGES ARE STRUCTURED:
# Image files are stored in a subfolder called "Cell_LineX_dashboard/".
# Brightfield images are named: CompoundName_DoseuM.jpeg
# DAPI images are named:        CompoundName_DoseuM_DAPI.jpeg
#
# HOW THE UI IS BUILT:
# htmltools::tagList() groups multiple HTML tags into a single output.
# htmltools::tags$div() creates HTML <div> containers.
# htmltools::tags$img() creates HTML <img> elements.
# CSS flexbox (display:flex) is used for layout throughout:
#   - The outer div is a horizontal row (image columns side by side)
#   - The left div has two images stacked vertically (DMSO BF + DAPI)
#   - The right div has two rows, each with three images side by side
#
# lapply() is used to generate the three repeated image slots per row.
# It takes a list of configurations (label text + img element ID) and 
# applies the same function to each, creating three identical-structured 
# <div> elements without repeating the HTML code manually.
#
# HOW THE DYNAMIC UPDATES WORK (JavaScript):
# The <script> block at the bottom defines a global function 
# window.updateCompoundImages(compound) that:
#   1. Constructs the expected image file path from the compound name and dose
#   2. Updates the src attribute of each <img> element to load the new image
#   3. Sets an onerror handler so if an image file doesn't exist, a 
#      "missing image" placeholder is shown instead of a broken image icon
#
# sprintf() embeds the first compound name into the JS at knit time 
# so the images load immediately on page open without waiting for a 
# table click (the %s is replaced with first_compound).
#
# The compound table's onRender() callback (above) calls 
# window.updateCompoundImages() whenever a row is clicked or the 
# table is redrawn, keeping the image panel in sync with the table.
# ============================================================

doses <- c("0.1", "1", "10")
first_compound <- disp_data_comp$Compound_Name[1]   # Used to pre-load images on page open

htmltools::tagList(
  
  # --- Outer container: full image panel, fixed height, horizontal layout ---
  htmltools::tags$div(
    style = "height:280px; overflow:hidden; display:flex; gap:12px;",

    # ===== LEFT SECTION: Static DMSO control images =====
    # Fixed width of 25%; two images stacked vertically with flex-direction:column
    htmltools::tags$div(
      style = "width:25%; display:flex; flex-direction:column; gap:12px; overflow:hidden;",

      # DMSO brightfield image (static — same file always loaded)
      htmltools::tags$div(
        style = "flex:1; display:flex; flex-direction:column; overflow:hidden; text-align:center;",
        htmltools::tags$div("DMSO – Brightfield", style = "font-weight:600; margin-bottom:4px;"),
        htmltools::tags$img(
          src   = "Cell_LineX_dashboard/DMSO_0.2%.jpeg",
          style = "flex:1; max-width:100%; max-height:100%; object-fit:contain;"
          # object-fit:contain scales the image to fit without cropping
        )
      ),

      # DMSO DAPI image (static)
      htmltools::tags$div(
        style = "flex:1; display:flex; flex-direction:column; overflow:hidden; text-align:center;",
        htmltools::tags$div("DMSO – DAPI", style = "font-weight:600; margin-bottom:4px;"),
        htmltools::tags$img(
          src   = "Cell_LineX_dashboard/DMSO_0.2%_DAPI.jpeg",
          style = "flex:1; max-width:100%; max-height:100%; object-fit:contain;"
        )
      )
    ),

    # ===== RIGHT SECTION: Dynamic compound images =====
    # Takes the remaining space (flex:1); two horizontal rows (BF + DAPI)
    htmltools::tags$div(
      style = "flex:1; display:flex; flex-direction:column; gap:12px; overflow:hidden;",

      # --- Row 1: Brightfield images at 0.1µM, 1µM, 10µM ---
      # lapply() generates three <div> blocks from the list of label/ID pairs,
      # avoiding repetition. Each list element contains: label text, img element ID.
      htmltools::tags$div(
        style = "flex:1; display:flex; gap:12px; overflow:hidden;",
        lapply(
          list(
            list("0.1 µM – Brightfield", "compound-image-bf-1"),
            list("1 µM – Brightfield",   "compound-image-bf-2"),
            list("10 µM – Brightfield",  "compound-image-bf-3")
          ),
          function(x)
            htmltools::tags$div(
              style = "width:33%; display:flex; flex-direction:column; overflow:hidden; text-align:center;",
              htmltools::tags$div(x[[1]], style = "font-weight:600; margin-bottom:4px;"),  # Label
              htmltools::tags$img(
                id    = x[[2]],   # Unique ID targeted by the JS below to update src
                style = "flex:1; max-width:100%; max-height:100%; object-fit:contain;"
              )
            )
        )
      ),

      # --- Row 2: DAPI images at 0.1µM, 1µM, 10µM ---
      htmltools::tags$div(
        style = "flex:1; display:flex; gap:12px; overflow:hidden;",
        lapply(
          list(
            list("0.1 µM – DAPI", "compound-image-dapi-1"),
            list("1 µM – DAPI",   "compound-image-dapi-2"),
            list("10 µM – DAPI",  "compound-image-dapi-3")
          ),
          function(x)
            htmltools::tags$div(
              style = "width:33%; display:flex; flex-direction:column; overflow:hidden; text-align:center;",
              htmltools::tags$div(x[[1]], style = "font-weight:600; margin-bottom:4px;"),
              htmltools::tags$img(
                id    = x[[2]],
                style = "flex:1; max-width:100%; max-height:100%; object-fit:contain;"
              )
            )
        )
      )
    )
  ),

  # ===== JAVASCRIPT: Global image update function =====
  # sprintf() is an R function that works like a template — the %s placeholder 
  # is replaced with the value of first_compound at knit time. This embeds the 
  # first compound name directly into the JavaScript string so the images load 
  # immediately when the page opens, without waiting for a user interaction.
  htmltools::tags$script(
    htmltools::HTML(sprintf("
      (function() {
        // Path to the placeholder shown when an image file is missing
        var placeholder = 'Cell_LineX_dashboard/missing_image.jpeg';
        // Folder where all compound images are stored
        var base  = 'Cell_LineX_dashboard/';
        // The three dose levels (must match the actual filename format)
        var doses = ['0.1', '1', '10'];

        // --- Helper: update one row of images (brightfield or DAPI) ---
        // compound: the compound name (used to build the file path)
        // ids:      array of HTML img element IDs to update
        // suffix:   '' for brightfield, '_DAPI' for DAPI images
        function updateRow(compound, ids, suffix) {
          doses.forEach(function(dose, i) {
            var img = document.getElementById(ids[i]);
            if (!img) return;

            // Construct the expected file path for this compound, dose, and image type
            var src = base + compound + '_' + dose + 'uM' + suffix + '.jpeg';

            // If the file doesn't exist, load the placeholder image instead
            // 'this.onerror = null' prevents infinite error loops
            img.onerror = function() {
              this.onerror = null;
              this.src = placeholder;
              this.title = compound + ' – ' + dose + ' µM (image missing)';
            };

            // Set the image source — the browser will fetch and display the file
            img.src = src;
            img.title = compound + ' – ' + dose + ' µM';
          });
        }

        // --- Global function called by the compound table's onRender() callback ---
        // This is attached to 'window' (the global browser scope) so it can be 
        // called from any other JavaScript on the page, including the table's 
        // onRender() callback defined in the compound table chunk above.
        window.updateCompoundImages = function(compound) {
          if (!compound) return;

          // Update brightfield row (no suffix)
          updateRow(
            compound,
            ['compound-image-bf-1', 'compound-image-bf-2', 'compound-image-bf-3'],
            ''
          );

          // Update DAPI row ('_DAPI' suffix in filename)
          updateRow(
            compound,
            ['compound-image-dapi-1', 'compound-image-dapi-2', 'compound-image-dapi-3'],
            '_DAPI'
          );
        };

        // Pre-load images for the first compound on page open.
        // The compound name is injected by R's sprintf() at knit time.
        window.updateCompoundImages('%s');
      })();
    ", first_compound))
  )
)
```
